---
title: "The role of osmorespiratory compromise in hypoxia tolerance of the purportedly oxyconforming teleost Galaxias maculatus"
author: "Jake Martin"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  # pdf_document:
  #   toc: true
  #   latex_engine: xelatex
  #   number_sections: true
  #   pandoc_args: ["--variable=documentclass:article"]
  html_document:
    code_download: true
    code_folding: hide
    depth: 4
    number_sections: no
    theme:  cosmo
    toc: yes
    toc_float: yes
    toc_depth: 4
knit: |
  (function(input, ...) {
    output_path <- rmarkdown::render(
      input,
      output_file = 'index.html',
      envir = globalenv()
    )
    file.copy(output_path, "supplementary-file-1.html", overwrite = TRUE)
    output_path
  })
---

[![License: CC BY 4.0](https://img.shields.io/badge/License-CC%20BY%204.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/) 

<!-- Comment out the license when knitting to PDF -->

<!------------------------------->

# 📕 READ ME

<!------------------------------->

**🚧 UNPUBLISHED AND ONGOING example 🚧**
The data in this script are currently **unpublished**, and the **analysis is ongoing**. Some sections may be incomplete. This script has been made publicly available to facilitate sharing with collaborators and colleagues who are interested in the progress of the project and to ensure transparency.

**SUMMARY**
This R code estimates the relationship between oxygen consumption (MO₂) and dissolved oxygen saturation (DO) in the Common Galaxias (*Galaxias maculatus*). It also estimates the critical oxygen saturation value for aerobic metabolism O<sub>2crit</sub>, commonly defined as the threshold below which the oxygen consumption rate can no longer be sustained. The associated article is titled *"The role of osmorespiratory compromise in hypoxia tolerance of the purportedly oxyconforming teleost Galaxias maculatus."* If you are reading the HTML version of this script, click the `Code` button in the top right to download the `.Rmd` file.

**AIM**
The article aims to test whether *Galaxias maculatus* can maintain oxygen consumption (MO₂) as ambient DO declines and, if so, at what level it reaches the critical oxygen saturation value for aerobic metabolism (O<sub>2crit</sub>).

**AUTHORS**
Timothy D. Clark ^[a]^
Maryane Gradito ^[a]^
Elizabeth C. Hoots ^[a]^
Luis L. Kuchenmüller ^[a]^
Jake M. Martin ^[a,b]^

**AFFILIATIONS**
[a] School of Life and Environmental Sciences, Deakin University, Geelong, VIC, Australia\
[b] Department of Wildlife, Fish and Environmental Studies, Swedish University of Agricultural Sciences, Umeå, Sweden\

**CONTRIBUTOR ROLES**
🚧 ***To be added*** 🚧
*Based on the Contributor Roles Taxonomy (CRediT)*

**DISCLAIMER**
I (Jake Martin) am dyslexic. I have made an effort to review the script for grammatical errors, but some will likely remain. I apologise. Please feel free to reach out using the contact details below if anything is unclear.

<!------------------------------->

# 📧 Contact

<!------------------------------->

**Jake M. Martin**

📧 **Email:** [jake.martin\@deakin.edu.au](mailto:jake.martin@deakin.edu.au)

📧 **Alt Email:** [jake.martin.research\@gmail.com](mailto:jake.martin.research@gmail.com)

🌐 **Web:** [jakemartin.org](https://jakemartin.org/)

🐙 **GitHub**: [JakeMartinResearch](https://github.com/JakeMartinResearch)

<!------------------------------->

# 📑 Sharing/accessing and citing

<!------------------------------->

1.  **Licenses/restrictions placed on the data:** CC-BY 4.0

2.  **Link to the associated publication:**\
    🚧 ***To be added*** 🚧

3.  **Recommended citation for this data:**\

Clark, T. D., Gradito, M., Hoots, E. C., Kuchenmüller, L. L., & Martin, J. M. (2025, May 28). The role of osmorespiratory compromise in hypoxia tolerance of the purportedly oxyconforming teleost Galaxias maculatus. https://doi.org/10.17605/OSF.IO/GFXCA

<!------------------------------->

# 📦 Required packages and knit settings

<!------------------------------->

A list of the required packages to run the script.

```{r setup}

# ----- Packages ------
if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")

# ---- List of required packages ----
pkgs <- c(
  # ----- Data Visualisation -----
  "ggthemes", "bayesplot", "gt", "gtsummary", "plotly", "qqplotr", "gridExtra",
  
  # ----- Tidy Data and Wrangling -----
  "tidyverse", "janitor", "readxl", "broom.mixed", "data.table", "hms", "devtools",
  "mclust",
  
  # ----- Modelling and Statistical Analysis -----
  "brms", "rstan", "marginaleffects", "performance", "emmeans",
  "tidybayes", "respirometry", "future"
)

# ---- Install and load all packages using pacman ----
suppressPackageStartupMessages(
  pacman::p_load(char = pkgs, install = TRUE)
)

#-------kniter seetting-----------

knitr::opts_chunk$set(
  echo = FALSE, # only for PDF knit
  message = FALSE,
  warning = FALSE, # no warnings
  cache = TRUE,# cacheing to save time when kniting
  tidy = TRUE,
  fig.align = "center" 
)
```


<!------------------------------->

# 🔧 Custom functions

<!------------------------------->

Here are some custom functions used within this script.

`bayes_incremental_regression_by_id()`: A custom function to build Bayesian incremental regressions. It is designed to run a list of subgroup models (IDs) in parallel using 4 cores. The function uses `brm()` with a Gaussian error distribution.

**Use**:\
The function accepts the following arguments:\
- `id_i`: A grouping factor or ID used to filter the data for each regression. If none is provided, the function uses the entire dataset.\
- `id_name`: The column name (as a character string) corresponding to the grouping factor in the data frame. If not provided, the function uses all data.\
- `data`: The data frame containing the variables for analysis.\
- `predictor`: The predictor variable of interest.\
- `response`: The response variable of interest.\
- `seed_number`: A random seed value for model reproducibility.\
- `save_models`: A logical argument indicating whether to save the model outputs (`TRUE` or `FALSE`).\
- `mod_output_wd`: The output directory where model `.rds` files should be saved (used only if `save_models = TRUE`).

```{r}
# Instead we could also use a distributional regression approach, by specifically modelling the variance by DO (e.g. sigma ~ DO). Weighting may not be required in this case, I don't think higher density of values in a given space will effect Bayesian estimates like it does in frequentist models. See discourse https://discourse.mc-stan.org/t/weights-in-brm/4278   

bayes_incremental_regression_by_id <- function(id_i, id_name, data, predictor, response, seed_number, save_models, mod_output_wd) {
  # Initiate an empty list to store models
  models <- list()
  
  # Check if id_name is missing, NULL, or blank, and assign NA if so
  if (missing(id_name) || is.null(id_name) || id_name == "") {
    id_name <- NA
  }
  
  # Check if id_i is missing, NULL, or blank, and assign NA if so
  if (missing(id_i) || is.null(id_i) || id_i == "") {
    id_name <- NA
  }
  
  # Filter data for the current ID if id_name is given as a factor or character and id_i is defined
  df_i <- data %>%
    dplyr::filter(
      if (!is.na(id_i) && (is.factor(data[[id_name]]) || is.character(data[[id_name]]))) {
        !!rlang::sym(id_name) == id_i
      } else {
        TRUE
      }
    )
  
  # Dynamically create formulas
  formula_lm_0 <- reformulate("1", response)
  formula_lm_1 <- reformulate(predictor, response)
  formula_lm_2 <- reformulate(sprintf("poly(%s, 2)", predictor), response)
  formula_lm_3 <- reformulate(sprintf("poly(%s, 3)", predictor), response)
  
  # Fit and store models in the list
  models[[paste0(id_i, "_lm_0")]] <- brm(
    bf(formula_lm_0, family = gaussian()), 
    data = df_i, cores = 4, seed = seed_number, save_pars = save_pars(all = save_models),
    sample_prior = FALSE, silent = TRUE, file = paste0(mod_output_wd, "/", id_i, "_lm_0")
  )
  
  models[[paste0(id_i, "_lm_1")]] <- brm(
    bf(formula_lm_1, family = gaussian()), 
    data = df_i, cores = 4, seed = seed_number, save_pars = save_pars(all = save_models),
    sample_prior = FALSE, silent = TRUE, file = paste0(mod_output_wd, "/", id_i, "_lm_1")
  )
  
  models[[paste0(id_i, "_lm_2")]] <- brm(
    bf(formula_lm_2, family = gaussian()), 
    data = df_i, cores = 4, seed = seed_number, save_pars = save_pars(all = save_models),
    sample_prior = FALSE, silent = TRUE, file = paste0(mod_output_wd, "/", id_i, "_lm_2")
  )
  
  models[[paste0(id_i, "_lm_3")]] <- brm(
    bf(formula_lm_3, family = gaussian()), 
    data = df_i, cores = 4, seed = 143019, save_pars = save_pars(all = save_models),
    sample_prior = FALSE, silent = TRUE, file = paste0(mod_output_wd, "/", id_i, "_lm_3")
  )
  
  # Return the list of models for the current ID
  return(models)
}
```

`load_rds9()`: A custom function to load all rds models in a directory and store in a list. Takes a directory with `.rds` files

```{r}
load_rds <- function(model_dw) {
  # List all .rds files in the directory
  model_file_list <- list.files(path = model_dw, pattern = "\\.rds$", full.names = TRUE)
  
  # Strip extensions and get names
  model_names <- tools::file_path_sans_ext(basename(model_file_list))
  
  # Read all .rds files into a named list
  model_store_list <- setNames(lapply(model_file_list, readRDS), model_names)
  
  return(model_store_list)
}
```

`incremental_regression_bayes_fits()`: A custom function for pulling model fits, loo and r2 using `loo()` and `bayes_R2()`, respectively. Takes a list of `brm` models.

```{r}
# Define Function to Process the data for each ID
incremental_regression_bayes_fits <- function(models) {
  
  loo_results_list <- list()
  
  # Iterate over the names of the models
  for (mod_name in names(models)) {
    # Extract the model
    mod_i <- models[[mod_name]]
    
    # Compute LOO results
    mod_loo_results_i <- loo::loo(mod_i)
    
    # Extract relevant LOO metrics
    elpd_loo_i <- mod_loo_results_i$elpd_loo
    p_loo_i <- mod_loo_results_i$p_loo
    looic_i <- mod_loo_results_i$looic
    
    # Create a data frame with metrics
    df_i <- data.frame(
      elpd_loo = elpd_loo_i,
      p_loo = p_loo_i,
      looic = looic_i,
      model = mod_name
    )
    
    est_i <- tidy(mod_i, effects = "fixed", conf.int = TRUE) %>% 
      dplyr::select(term, estimate, conf.low, conf.high) %>% 
      tidyr::pivot_wider(
        names_from = term,                # Use `term` as column names
        values_from = c(estimate, conf.low, conf.high),  # Values to pivot
        names_sep = "_"                   # Add a separator to column names
      )
    
    df_i <- cbind(df_i, est_i)
    
    # Store the data frame in the list
    loo_results_list[[mod_name]] <- df_i
  }
  
  # Combind 
  loo_results_combined <- bind_rows(loo_results_list)
  
  # Get R2 
  r2_results <- map_dfr(models, ~ as.data.frame(bayes_R2(.x)), .id = "model") %>%
    tibble::remove_rownames()
  
  # Combind R2 and loo results 
  model_fit_df <- dplyr::full_join(loo_results_combined, r2_results, by = "model") %>% 
    dplyr::select(model, everything()) %>% 
    dplyr::rename(r2 = Estimate,
                  r2_error = Est.Error,
                  r2_q2.5 = Q2.5,
                  r2_q97.5 = Q97.5) %>% 
    dplyr::mutate(id = sub("_(lm_\\d+)$", "", model),
                  model_type = sub("^.*_(lm_\\d+)$", "\\1", model))
  
  return(model_fit_df)
}
```

`bayes_mod_predictions()`: This function extracts the predicted values using `fitted()` from a list of models and combines them with the original data file used for the model. These are the **posterior mean fitted values** (i.e. the expected value of the response variable given the predictor variables and the estimated posterior distributions of the parameters) for each observation in the dataset, along with **95% credible intervals**.

```{r}
bayes_mod_predictions <- function(models, original_data) {
  
  prediction_list <- list()
  
  for (mod_name in names(models)) {
    # Extract mod
    mod_i <- models[[mod_name]]
    
    # Make mode predictions
    model_predictions_i <- as.data.frame(fitted(mod_i, summary = TRUE)) %>% 
      dplyr::mutate(model = mod_name,
                    id = sub("_(lm_\\d+)$", "", mod_name),
                    model_type = sub("^.*_(lm_\\d+)$", "\\1", mod_name)) %>% 
      dplyr::rename(pred_lower = Q2.5, pred_upper = Q97.5, predicted = Estimate, pred_error = Est.Error) %>% 
      dplyr::select(model, everything())
    
    id_i <- model_predictions_i$id[1]
    
    original_data_i <- original_data %>% 
      dplyr::filter(id == id_i) %>% 
      dplyr::select(-id)
    
    model_predictions_original_i <- cbind(model_predictions_i, original_data_i)
    
    prediction_list[[mod_name]] <- model_predictions_original_i
  }
  predictions_df <- bind_rows(prediction_list) 
  return(predictions_df)
}
```

`calcSMR()`: **authored by Denis Chabot** used to estimate SMR with several different methods Claireaux and Chabot (2016) ^[1]^

^[1]^ Claireaux, G. and Chabot, D. (2016) Responses by fishes to environmental hypoxia: integration through Fry's concept of aerobic metabolic scope. Journal of Fish Biology <https://doi.org/10.1111/jfb.12833>

```{r}
calcSMR = function(Y, q=c(0.1,0.15,0.2,0.25,0.3), G=1:4){
	u = sort(Y)
	the.Mclust <- Mclust(Y,  G=G)
	cl <- the.Mclust$classification
	# sometimes, the class containing SMR is not called 1
	# the following presumes that when class 1 contains > 10% of cases, 
	# it contains SMR, otherwise we take class 2
	cl2 <- as.data.frame(table(cl))
	cl2$cl <- as.numeric(levels(cl2$cl))
	valid <- cl2$Freq>=0.1*length(time)  
	the.cl <- min(cl2$cl[valid])
	left.distr <- Y[the.Mclust$classification==the.cl]
	mlnd = the.Mclust$parameters$mean[the.cl]
	CVmlnd = sd(left.distr)/mlnd * 100
	quant=quantile(Y, q)
	low10=mean(u[1:10])
	low10pc = mean(u[6:(5 + round(0.1*(length(u)-5)))])
	# remove 5 outliers, keep lowest 10% of the rest, average
	# Herrmann & Enders 2000
	return(list(mlnd=mlnd, quant=quant, low10=low10, low10pc=low10pc,
		      cl=cl, CVmlnd=CVmlnd))
}
```

`calcO2crit()`: **authored by Denis Chabot** used to estimate O2crit (Pcript). Claireaux and Chabot (2016) ^[1]^

***Note: O2 is assumed to be in percentage of dissolved oxygen (DO)***

```{r}
calcO2crit <- function(Data, SMR, lowestMO2 = NA, gapLimit = 4,
max.nb.MO2.for.reg = 20)
{
# AUTHOR: Denis Chabot, Institut Maurice-Lamontagne, DFO, Canada
# first version written in June 2009
# last updated in January 2015
method = "LS_reg" # will become "through_origin" if intercept is > 0
if(is.na(lowestMO2)) lowestMO2 = quantile(Data$MO2[Data$DO >= 80], p=0.05)
# Step 1: identify points where MO2 is proportional to DO
geqSMR = Data$MO2 >= lowestMO2
pivotDO = min(Data$DO[geqSMR])
lethal = Data$DO < pivotDO
N_under_SMR = sum(lethal) # points available for regression?
final_N_under_SMR = lethal # some points may be removed at Step 4
lastMO2reg = Data$MO2[Data$DO == pivotDO] # last MO2 when regulating
if(N_under_SMR > 1) theMod = lm(MO2~DO, data=Data[lethal,])
# Step 2, add one or more point at or above SMR
# 2A, when there are fewer than 3 valid points to calculate a regression
if(N_under_SMR < 3){
missing = 3 - sum(lethal)
not.lethal = Data$DO[geqSMR]
DOlimit = max(sort(not.lethal)[1:missing]) # highest DO acceptable
# to reach a N of 3
addedPoints = Data$DO <= DOlimit
lethal = lethal | addedPoints
theMod = lm(MO2~DO, data=Data[lethal,])
}
# 2B, add pivotDO to the fit when Step 1 yielded 3 or more values?
if(N_under_SMR >= 3){
lethalB = Data$DO <= pivotDO # has one more value than "lethal"
regA = theMod
regB = lm(MO2~DO, data=Data[lethalB,])
large_slope_drop = (coef(regA)[2]/coef(regB)[2]) > 1.1 # arbitrary
large_DO_gap = (max(Data$DO[lethalB]) - max(Data$DO[lethal])) > gapLimit
tooSmallMO2 = lastMO2reg < SMR
if(!large_slope_drop & !large_DO_gap & !tooSmallMO2) {
lethal = lethalB
theMod = regB
} # otherwise we do not accept the additional point
}
# Step 3
# if the user wants to limit the number of points in the regression
if(!is.na(max.nb.MO2.for.reg) & sum(lethal)>max.nb.MO2.for.reg){
Ranks = rank(Data$DO)
lethal = Ranks <= max.nb.MO2.for.reg
theMod = lm(MO2~DO, data=Data[lethal,])
final_N_under_SMR = max.nb.MO2.for.reg
}
# Step 4
predMO2 = as.numeric(predict(theMod, data.frame(DO=Data$DO)))
Data$delta = (Data$MO2-predMO2)/predMO2 * 100 # residuals set to zero
# when below pivotDO
Data$delta[Data$DO < pivotDO | lethal] = 0
tol = 0 # any positive residual is unacceptable
HighValues = Data$delta > tol
Ranks = rank(-1*Data$delta)
HighMO2 = HighValues & Ranks == min(Ranks) # keep largest residual
if (sum(HighValues) > 0) {
nblethal = sum(lethal)
Data$W = NA
Data$W[lethal]=1/nblethal
Data$W[HighMO2] = 1
theMod = lm(MO2~DO, weight=W, data=Data[lethal | HighMO2,])
# This new regression is always an improvement, but there can still
# be points above the line, so we repeat
predMO2_2 = as.numeric(predict(theMod, data.frame(DO=Data$DO)))
Data$delta2 = (Data$MO2-predMO2_2)/predMO2_2 * 100
Data$delta2[Data$DO < pivotDO] = 0
tol = Data$delta2[HighMO2]
HighValues2 = Data$delta2 > tol
if(sum(HighValues2)>0){
Ranks2 = rank(-1*Data$delta2)
HighMO2_2 = HighValues2 & Ranks2 == 1 # keep the largest residual
nblethal = sum(lethal)
Data$W = NA
Data$W[lethal]=1/nblethal
Data$W[HighMO2_2] = 1
theMod2 = lm(MO2~DO, weight=W, data=Data[lethal | HighMO2_2,])
# is new slope steeper than the old one?
if(theMod2$coef[2] > theMod$coef[2]) {
theMod = theMod2
HighMO2 = HighMO2_2
}
} # end second search for high value
} # end first search for high value
Coef = coefficients(theMod)
#Step 5, check for positive intercept
AboveOrigin = Coef[1] > 0
# if it is, we use a regression that goes through the origin
if (AboveOrigin){
theMod = lm(MO2~DO -1, data=Data[lethal,])
Coef = c(0, coefficients(theMod)) # need to add the intercept (0)
# manually to have a pair of coefficients
method = "through_origin"
HighMO2 = rep(FALSE, nrow(Data)) # did not use the additional value
# from Step 4
}
po2crit = as.numeric(round((SMR - Coef[1]) / Coef[2], 1))
sum_mod = summary(theMod)
anov_mod = anova(theMod)
O2CRIT = list(o2crit=po2crit, SMR=SMR, Nb_MO2_conforming = N_under_SMR,
Nb_MO2_conf_used = final_N_under_SMR,
High_MO2_required = sum(HighMO2) == 1, origData=Data,
Method=method, mod=theMod, r2 = sum_mod$r.squared,
P = anov_mod$"Pr(>F)", lethalPoints = which(lethal),
AddedPoints = which(HighMO2))
} # end function
```

`plotO2crit()`: **authored by Denis Chabot**, used to plot the modes used for the `calcO2crit()` function. Claireaux and Chabot (2016) ^[1]^

***Note: I added abline(h=lowestMO2, col="pink") so that I could visualise the lowestMO2 position***

```{r}
plotO2crit <- function(o2critobj, plotID="",
Xlab="Dissolved oxygen (% sat.)", Ylab="dotitalumol",
smr.cex=0.9, o2crit.cex=0.9, plotID.cex=1.2,
Transparency=T,...)
{
# AUTHOR: Denis Chabot, Institut Maurice-Lamontagne, DFO, Canada
# first version written in June 2009
# last updated 2015-02-09
# for R plotting devices that do not support transparency
# (e.g., postscript), set Transparency to FALSE
smr = o2critobj$SMR
if(Ylab %in% c("dotitalumol", "italumol", "dotumol", "umol",
"dotitalmg", "italmg", "dotmg", "mg")) {
switch(Ylab,
dotitalumol = {
mo2.lab = expression(paste(italic(dot(M))[O[2]], " (",mu,"mol ", O[2],
" ", min^-1, " ", kg^-1, ")"))
},
italumol = {
mo2.lab = expression(paste(italic(M)[O[2]], " (",mu,"mol ", O[2], " ",
min^-1, " ", kg^-1, ")"))
},
dotumol = {
mo2.lab = expression(paste(dot(M)[O[2]], " (",mu,"mol ", O[2], " ",
min^-1, " ", kg^-1, ")"))
},
umol = {
mo2.lab = expression(paste(M[O[2]], " (",mu,"mol ", O[2], " ", min^-1,
" ", kg^-1, ")"))
},
dotitalmg = {
mo2.lab = expression(paste(italic(dot(M))[O[2]], " (mg ", O[2], " ",
h^-1, " ", kg^-1, ")"))
},
italmg = {
mo2.lab = expression(paste(italic(M)[O[2]], " (mg ", O[2], " ",
h^-1, " ", kg^-1, ")"))
},
dotmg = {
mo2.lab = expression(paste(dot(M)[O[2]], " (mg ", O[2], " ", h^-1, " ",
kg^-1, ")"))
},
mg = {
mo2.lab = expression(paste(M[O[2]], " (mg ", O[2], " ", h^-1, " ",
kg^-1, ")"))
}
)
} else mo2.lab=Ylab
if(Transparency) {Col=c(rgb(0,0,0,0.7), "red", "orange")
} else {Col=c(grey(0.3), "red", "orange")}
Data=o2critobj$origData
lowestMO2 = quantile(Data$MO2[Data$DO >= 80], p=0.05) # I added this
Data$Color = Col[1]
Data$Color[o2critobj$lethalPoints] = Col[2]
Data$Color[o2critobj$AddedPoints] = Col[3]
# ordinary LS regression without added points: blue line, red symbols
# ordinary LS regression with added points: blue line, red & orange symbols
# regression through origin: green dotted line, red symbols
line.color = ifelse(o2critobj$Method=="LS_reg", "blue", "darkgreen")
line.type = ifelse(o2critobj$Method=="LS_reg", 1, 3)
limX = c(0, max(Data$DO))
limY = c(0, max(Data$MO2))
plot(MO2~DO, data=Data, xlim=limX, ylim=limY, col=Data$Color, xlab=Xlab,
ylab=mo2.lab, ...)
coord <- par("usr")
if(plotID != ""){
text(0, coord[4], plotID, cex=plotID.cex, adj=c(0,1.2))
}
abline(h=lowestMO2, col="pink") # I added this
abline(h=smr, col="orange")
text(coord[1], smr, "SMR", adj=c(-0.1,1.3), cex=smr.cex)
text(coord[1], smr, round(smr,1), adj=c(-0.1,-0.3), cex=smr.cex)
if(!is.na(o2critobj$o2crit)) {
abline(o2critobj$mod, col=line.color, lty=line.type)
segments(o2critobj$o2crit, smr, o2critobj$o2crit, coord[3],
col=line.color, lwd=1)
text(x=o2critobj$o2crit, y=0, o2critobj$o2crit, col=line.color,
cex=o2crit.cex, adj=c(-0.1,0.5))
}
} # end of function
```

<!------------------------------->

# 📂 Directories

<!------------------------------->

## Input

**📥 input_data_wd**: Directory for the metadata

```{r}
wd <- getwd()
input_data_wd <- paste0(wd, "./input-data") # creates a variable with the name of the wd we want to use
```

**📥 mod_data_wd**: Directory for model output data estimated slopes

```{r}
mod_data_wd <- paste0(wd, "./mod-data")
```

## Output

**📤 output_fig_wd**: this is where we will put the figures

```{r}
output_fig_wd <- paste0(wd, "./output-fig")
if (!dir.exists("output-fig")) {dir.create("output-fig")}
```

**📤 output_mods_wd**: this is where we will put the models

```{r}
output_mods_wd <- paste0(wd, "./output-mod")
if (!dir.exists("output-mod")) {dir.create("output-mod")}
```

<!------------------------------->

# 💿 Data

<!------------------------------->

## Slopes (ṀO)

**💿 slope_df**: We have imported the slopes extracted in LabChart during each phase of the experiment

```{r}
 setwd(input_data_wd)

data_file <- "labchart-slope-data.xlsx"

# 
# # Get the names of all sheets in the Excel file
sheet_names <- readxl::excel_sheets(data_file)
all_trials_select <- c("start_date", "order", "phase", "cycle", "date", "time")
slope_list <- list()

for (sheet in sheet_names) {

  df <- read_excel(data_file, sheet = sheet) %>% 
  dplyr::rename_with(tolower)
  
a_name <- paste0("a_", tolower(sheet))
a_df <- df %>%
  dplyr::select(starts_with('a'), all_trials_select) %>% 
  dplyr::rename(temp = a_temp) %>% 
  dplyr::mutate(across(starts_with('a'), as.numeric)) %>% 
  pivot_longer(
    cols = starts_with('a'), # Select all columns to pivot
    names_to = c("chamber_id", ".value"), # Separate column names into 'id' and other variables
    names_sep = "_"
  ) %>%
  dplyr::mutate(respirometer_group = "a") # Add a new column with a fixed value

slope_list[[a_name]]<- a_df

b_name <- paste0("b_", tolower(sheet))
b_df <- df %>% 
  dplyr::select(starts_with('b'), all_trials_select) %>% 
  dplyr::rename(temp = b_temp) %>% 
  dplyr::mutate(across(starts_with('b'), as.numeric)) %>% 
  pivot_longer(
    cols = starts_with('b'), # Select all columns to pivot
    names_to = c("chamber_id", ".value"), # Separate column names into 'id' and other variables
    names_sep = "_"
  ) %>% 
    dplyr::mutate(respirometer_group = "b")

slope_list[[b_name]] <- b_df

c_name <- paste0("c_", tolower(sheet))
c_df <- df %>% 
  dplyr::select(starts_with('c'), all_trials_select) %>% 
  dplyr::rename(temp = c_temp,
                i_cycle = cycle) %>% 
  dplyr::mutate(across(starts_with('c'), as.numeric)) %>%
  pivot_longer(
    cols = starts_with('c'), # Select all columns to pivot
    names_to = c("chamber_id", ".value"), # Separate column names into 'id' and other variables
    names_sep = "_"
  ) %>% 
    dplyr::mutate(respirometer_group = "c") %>% 
  dplyr::rename(cycle = i_cycle)

slope_list[[c_name]] <- c_df

d_name <- paste0("d_", tolower(sheet))
d_df <- df %>% 
  dplyr::select(starts_with('d'), all_trials_select) %>% 
  dplyr::rename(temp = d_temp,
                i_date = date) %>% 
  dplyr::mutate(across(starts_with('d'), as.numeric)) %>%
  pivot_longer(
    cols = starts_with('d'), # Select all columns to pivot
    names_to = c("chamber_id", ".value"), # Separate column names into 'id' and other variables
    names_sep = "_"
  ) %>% 
    dplyr::mutate(respirometer_group = "d") %>% 
  dplyr::rename(date = i_date)

slope_list[[d_name]] <- d_df
}


slope_df <- bind_rows(slope_list) %>% 
  dplyr::mutate(resp_cat_date = paste0(respirometer_group, "_", start_date),
                chamber_n = str_extract(chamber_id, "\\d+"),
                id_prox = paste0(resp_cat_date, "_", chamber_n),
                time_hms = as_hms(time*3600),
                date_chr = format(date, "%d/%m/%Y")
                )
```

## Fish metadata

**💿 metadata**: This is the meta data for each chamber

*Note: We are also adding volume based on chamber type.*

```{r}
setwd(input_data_wd)

metadata <- read_excel("fish-meta.xlsx", na = "NA") %>% 
  dplyr::mutate(id_split = id) %>% 
  tidyr::separate(id_split, into = c("respirometer_group", "salinity_group", "start_date", "chamber"), sep = "_") %>% 
  dplyr::mutate(
      volume = dplyr::case_when(
        chamber_type == "L" ~ 0.300,
        chamber_type == "M_M" ~ 0.105,
        chamber_type == "M_NM" ~ 0.11,
        chamber_type == "S" ~ 0.058,
        chamber_type == "SM" ~ 0.075,
        chamber_type == "D3" ~ 0.055,
        TRUE ~ NA
      ),
      id_prox = paste0(respirometer_group, "_", start_date, "_", chamber),
      rel_size = mass/volume)
```

## Urbina, Glover, and Forster (2012)

**💿 urbina_et_al_2012**: This is the mean level data extracted from Urbina, Glover, and Forster (2012)^[2]^ Figure 1a. We used the metaDigitise package in R to extract the data ^[3]^.

```{r}
setwd(input_data_wd)

urbina_et_al_2012 <- read_excel("urbina-et-al-2012-fig1a.xlsx")
```

## O₂crit viusal inspection

**💿 o2crit_check**: This data frame includes the visual based assessment for O₂crit for all fish included in the analysis (*n* = 58). The inspections were made using figures genbrated by the `plotO2crit()` function, and can be viewed in `combined_chabot_plots.pdf`. The visual assessment was done independently by all authors, and the presence of an O₂crits allocated to 'yes', 'no', 'maybe'. If yes, or maybe, an estimate of the dissolved oxygen percentage was given. 

Directory to the figures   

📁 gmac-mo2-24/
└── 📂 output-fig/  
    └── 📂 model_chabot/ 
         └── 📄 combined_chabot_plots.pdf   

```{r}
setwd(input_data_wd)

o2crit_check <- read_excel("o2crit-visual-assessment.xlsx", na = "NA")
```

### Combinding metadata

Adding the meta data to the slopes data frame

```{r}
slope_df_2 <- slope_df %>% 
  dplyr::select(-start_date, -respirometer_group) %>% 
  left_join(metadata, by = "id_prox") %>% 
  dplyr::mutate(light_dark = if_else(time_hms >= as.hms("07:00:00") & time_hms < as.hms("19:00:00"), "light", "dark")) %>% 
  dplyr::arrange(id)
```

<!------------------------------->

# 🧹 Data tidy

<!------------------------------->

## Numbers

We have **64 fish** with MO₂ data

```{r}
n <- slope_df_2 %>% 
  dplyr::filter(chamber_condition == "fish") %>% 
  dplyr::distinct(id) %>% 
  nrow(.)

paste0("n = ", n)
```

With 48 from the 0 ppt and 48 from 9 ppt groups

```{r}
slope_df_2 %>% 
  dplyr::filter(chamber_condition == "fish") %>%
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe('n total' = length(unique(id))) %>% 
  gt() %>% 
  cols_label(
    salinity_group = "Salinity group"
  ) %>% 
  cols_align(
    align = "center", 
    columns = everything()
  )
```

## Fish size

Here we calculate the mean length and size of fish used in the experiment.

```{r}
mass_length <- slope_df_2 %>% 
  dplyr::filter(chamber_condition == "fish") %>%
  dplyr::group_by(id) %>% 
  dplyr::sample_n(1) %>% 
  dplyr::ungroup() %>% 
  dplyr::reframe(x_mass = round(mean(mass, na.rm = TRUE), 3),
                 min_mass = round(min(mass, na.rm = TRUE), 3),
                 max_mass = round(max(mass, na.rm = TRUE), 3),
                 x_length = round(mean(length, na.rm = TRUE), 2),
                 min_length = round(min(length, na.rm = TRUE), 2),
                 max_length = round(max(length, na.rm = TRUE), 2))

mass_mean <- mass_length %>% 
  pull(x_mass)

mass_min <- mass_length %>% 
  pull(min_mass)

mass_max <- mass_length %>% 
  pull(max_mass)

length_mean <- mass_length %>% 
  pull(x_length)

length_min <- mass_length %>% 
  pull(min_length)

length_max <- mass_length %>% 
  pull(max_length)

paste0("The mean mass of fish was ", mass_mean, " g (range: ", mass_min, "-", mass_max, ")",
       ", and the mean length was ", length_mean, " mm (range: ", length_min, "-", length_max, ")")
```

## Filtering trials

We will remove 6 trials which had errors. These are as follows:

-   a_0_25nov_3 fish died during trial
-   b_0_26nov_4 flat lined early
-   c_0_22nov_2 chamber was opened early
-   c_9_26nov_2 chamber was opened early
-   c_9_26nov_4 chamber was opened early
-   d_9_27nov_3 sensor was jumpy and end points were hard to confidently ID visually

```{r}
remove_trial_error <- c("a_0_25nov_3", "b_0_26nov_4", "c_0_22nov_2", "c_9_26nov_2", "c_9_26nov_4", "d_9_27nov_3")

slope_df_filter <- slope_df_2 %>% 
  dplyr::filter(!(id %in% remove_trial_error))
```

We now have **58 fish** with MO2 data

```{r}
n <- slope_df_filter %>% 
  dplyr::filter(chamber_condition == "fish") %>% 
  dplyr::distinct(id) %>% 
  nrow(.)

paste0("n = ", n)
```

With 30 in the 0 ppt group and 28 in the 9 ppt group

```{r}
slope_df_filter %>% 
  dplyr::filter(chamber_condition == "fish") %>%
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe('n total' = length(unique(id))) %>% 
  gt() %>% 
  cols_label(
    salinity_group = "Salinity group"
  ) %>% 
  cols_align(
    align = "center", 
    columns = everything()
  )
```

Mass and size of fish after removing 6 trials which had errors

```{r}
mass_length <- slope_df_filter %>% 
  dplyr::filter(chamber_condition == "fish") %>%
  dplyr::group_by(id) %>% 
  dplyr::sample_n(1) %>% 
  dplyr::ungroup() %>% 
  dplyr::reframe(x_mass = round(mean(mass, na.rm = TRUE), 3),
                 min_mass = round(min(mass, na.rm = TRUE), 3),
                 max_mass = round(max(mass, na.rm = TRUE), 3),
                 x_length = round(mean(length, na.rm = TRUE), 2),
                 min_length = round(min(length, na.rm = TRUE), 2),
                 max_length = round(max(length, na.rm = TRUE), 2))

mass_mean <- mass_length %>% 
  pull(x_mass)

mass_min <- mass_length %>% 
  pull(min_mass)

mass_max <- mass_length %>% 
  pull(max_mass)

length_mean <- mass_length %>% 
  pull(x_length)

length_min <- mass_length %>% 
  pull(min_length)

length_max <- mass_length %>% 
  pull(max_length)

paste0("The mean mass of fish was ", mass_mean, " g (range: ", mass_min, "-", mass_max, ")",
       ", and the mean length was ", length_mean, " mm (range: ", length_min, "-", length_max, ")")
```

Chamber size breakdown

```{r}
slope_df_filter %>% 
  dplyr::filter(chamber_condition == "fish") %>%
  dplyr::mutate(volume = if_else(volume == 0.055, 0.058, volume)) %>% 
  dplyr::group_by(volume) %>% 
  dplyr::reframe('n total' = length(unique(id)),
                 median_mass = median(mass, na.rm = TRUE),
                 sd_mass = sd(mass, na.rm = TRUE)) %>% 
  gt() %>% 
  cols_label(
    volume = "Chamber type",
    median_mass = "Median mass",
    sd_mass = "Standard deviation mass"
  ) %>% 
  cols_align(
    align = "center", 
    columns = everything()
  )
```

Numbers per trial approach

```{r}
phases <- c("50c", "100c")

slope_df_filter %>% 
  dplyr::filter(chamber_condition == "fish" & phase %in% phases) %>%
  dplyr::group_by(phase) %>% 
  dplyr::reframe('n total' = length(unique(id))) %>% 
  gt() %>% 
  cols_label(
    phase = "Type of trial",
  ) %>% 
  cols_align(
    align = "center", 
    columns = everything()
  )
```

## Filtering MO₂ estimates

Here we apply the following filters to the MO₂ data:

-   Remove the first 5 SMR cycles (burn in)
-   Remove all positive raw slopes
-   Remove all MO₂ calculated using less then 60 data points (5 min at 0.2 Hz)
-   Remove all MO₂ calculated if O₂ increases in a closed phase (i.e. trial has ended)

Check positive values for MO₂ before removing.

```{r}
slope_tidy_remove_flush <- slope_df_filter %>%
  dplyr::filter(phase != "smr", chamber_condition == "fish") %>% 
  dplyr::group_by(id) %>%
  dplyr::arrange(id, order) %>%  # Ensure the data is sorted within each group
  dplyr::mutate(o2_diff = if_else(row_number() == 1, 0, o2 - lag(o2)), # Calculate the difference in 'o2'
                o2_diff_cumsum = cumsum(o2_diff > 1)) %>%  # Checks first occurrence and sums 
  dplyr::filter(o2_diff_cumsum == 0) %>%  # Keep rows until the first jump > 1
  dplyr::ungroup() %>%
  dplyr::select(-o2_diff, -o2_diff_cumsum)

postive_slopes <- slope_tidy_remove_flush %>%
  dplyr::filter(chamber_condition == "fish" & mo2corr > 0)

n_postive_slopes <- nrow(postive_slopes)

n_slopes <-  nrow(slope_tidy_remove_flush)

list_postive_all <- postive_slopes %>% 
  dplyr::distinct(id) %>% 
  dplyr::pull(id)

print(paste0("There are ", length(list_postive_all), " fish with postive slopes. These fish are: ", paste(list_postive_all, collapse = ", "), ". For all estimated slopes (n = ", n_slopes, ") ", round((n_postive_slopes/n_slopes)*100,2) , "% were postive (n = ", n_postive_slopes, ")"))
```

Filtering the MO₂ data

```{r}
cycle_burn <- 0:4

slope_df_filter_1 <- slope_df_filter %>%
  dplyr::filter(!(cycle %in% cycle_burn) & 
                  mo2corr < 0 & 
                  n > 60 &
                  chamber_condition == "fish"
                )
  
# Filter out the end flush
slope_tidy_closed <- slope_df_filter_1 %>%
  dplyr::filter(phase != "smr") %>% 
  dplyr::group_by(id) %>%
  dplyr::arrange(id, order) %>%  # Ensure the data is sorted within each group
  dplyr::mutate(o2_diff = if_else(row_number() == 1, 0, o2 - lag(o2)), # Calculate the difference in 'o2'
                o2_diff_cumsum = cumsum(o2_diff > 1)) %>%  # Checks first occurrence and sums 
  dplyr::filter(o2_diff_cumsum == 0) %>%  # Keep rows until the first jump > 1
  dplyr::ungroup() %>%
  dplyr::select(-o2_diff, -o2_diff_cumsum)

slope_tidy_smr <- slope_df_filter_1 %>% 
  dplyr::filter(phase == "smr")

slope_df_filter_2 <- rbind(slope_tidy_smr, slope_tidy_closed) %>% 
  dplyr::arrange(id, order)
```

## Calculating SMR

We will calculate SMR using `calcSMR` function by Chabot, Steffensen and Farrell (2016)^[1]^. Specifically, we use mean of the lowest normal distribution (MLND) where CVmlnd \< 5.4 and the mean of the lower 20% quantile (q0.2) were CVmlnd \> 5.4. If CVmlnd is not calculated we have used q0.2.

```{r}
labchart_chabot_smr <- slope_df_filter_2 %>%
  dplyr::filter(phase == "smr")

# Extract distinct IDs
ids <- labchart_chabot_smr %>% 
  dplyr::distinct(id) %>% 
  dplyr::pull()

# Initialise an empty list to store SMR data
smr_list <- list()

# Process each ID
for (id_i in ids) {
  tryCatch({
    # Filter the data for the current ID
    df_i <- labchart_chabot_smr %>% 
      dplyr::filter(id == id_i) %>% 
      dplyr::mutate(abs_mo2corr = abs(mo2corr))
    
    # Calculate SMR results
    calcSMR_results <- calcSMR(df_i$abs_mo2corr)
    CVmlnd_i <- calcSMR_results$CVmlnd
    quant_i <- calcSMR_results$quant %>% as_tibble()
    quant_20per_i <- quant_i$value[3]
    mlnd_i <- calcSMR_results$mlnd
    smr_value <- if_else(CVmlnd_i < 5.4, mlnd_i, quant_20per_i)
    smr_type <- if_else(CVmlnd_i < 5.4, "mlnd", "quant_20per")
    smr_value <- if_else(is.na(smr_value), quant_20per_i, smr_value)
    smr_type <- if_else(is.na(smr_type), "quant_20per", smr_type)
    
    # Create a data frame for the current ID
    smr_df_i <- tibble::tibble(
      id = id_i,
      smr = smr_value,
      smr_est = smr_type
    )
    
  }, error = function(e) {
    # Handle errors by assigning NA values
    smr_df_i <- tibble::tibble(
      id = id_i,
      smr = NA,
      smr_est = NA
    )
  })
  
  # Append to the list
  smr_list[[id_i]] <- smr_df_i
}

# Combine all individual SMR data frames into one
smr_df <- bind_rows(smr_list) %>% 
  dplyr::rename(smr_chabot = smr,
                smr_chabot_method = smr_est)

slope_df_filter_3 <- slope_df_filter_2 %>%
  dplyr::left_join(., smr_df, by = "id")
```

## Transforming MO₂

Here we are transforming the MO₂ units. The resulting values are as follows:

-   **MO2** is absolute value of the background and leak corrected MO₂ slope from Labchart (mo2corr) times the net volume of the chamber (volume - fish mass), × 60, × 60, to achieve MO₂ as mg^-1^ O₂ h^-1^

-   **MO2_g** is MO2 divided by fish mass to achieve MO₂ as mg^-1^ O₂ g^-1^ h^-1^ (i.e. mass standardised)

-   **SMR** absolute value of the SMR estimates using methods described by Chabot, Steffensen and Farrell (2016)^[1]^ times the net volume of the chamber (volume - fish mass), × 60, × 60, to achieve SMR as mg^-1^ O₂ h^-1^)

-   **SMR_g** is SMR divided by fish mass to achieve SMR as mg^-1^ O₂ g^-1^ h^-1^ (i.e. mass standardised)

-   **DO** is dissolved oxygen percentage calculated from O₂ values (mg^-1^ L^-1^) using the recorded temperature, salinity, and a constant atmospheric pressure (Pa; 1013.25)

-   **o2_kpa** is the O₂ concentration in kilopascal (kpa). This is used to make a comparative figure only.

```{r}
# Combine back into one data frame
slope_tidy <- slope_df_filter_3 %>% 
    dplyr::mutate(DO = conv_o2(
                   o2 = o2,
                   from = "mg_per_l",
                   to = "percent_a.s.",
                   temp = temp, #C
                   sal = measured_salinity,
                   atm_pres = 1013.25),
                  o2_kpa = conv_o2(
                   o2 = o2,
                   from = "mg_per_l",
                   to = "kPa",
                   temp = temp, #C
                   sal = measured_salinity,
                   atm_pres = 1013.25),
                  net_volume = volume - mass/1000,
                  MO2 = abs(mo2corr)*net_volume*60*60,
                  MO2_g = MO2/mass,
                  SMR = abs(smr_chabot)*net_volume*60*60,
                  SMR_g = SMR/mass
                  )
```

Trial lengths

```{r}
slope_tidy %>% 
  dplyr::filter(phase != "smr") %>% 
  dplyr:: mutate(datetime = as.POSIXct(paste(date, time_hms), format = "%Y-%m-%d %H:%M:%S")) %>%
  dplyr::group_by(id) %>% 
  dplyr::reframe(
    phase = phase[1],
    start_time = min(datetime, na.rm = TRUE),
    end_time = max(datetime, na.rm = TRUE),
    time_diff = difftime(end_time, start_time, units = "hours")  # Change units as needed
  ) %>% 
  dplyr::group_by(phase) %>% 
  dplyr::reframe('Minimum duration' = round(min(time_diff),2),
                 'Maximum duration' = round(max(time_diff),2),
                 'Median duration' = round(median(time_diff),2)) %>% 
  dplyr::mutate(phase = if_else(phase == "100c", "Closed at 100% air saturation", "Closed at 75% and 50% air saturation")) %>% 
  dplyr::rename(Approach = phase) %>% 
  gt()
  
```

## Visualising MO2 corrections  

Let's check how data corrections change the appearance of out data 
MO2

Let's compare raw MO2 calculations (mo2), MO2 calculations after correcting background respiration (mo2_bground), and MO2 calculations after correcting for background respiration and oxygen diffusion dynamics (mo2corr). First we will need to make the mo2_bground vaule as it's not present in the datafile, and transform the data to long formate, so we can plot it more easly.


```{r}
slope_corr_comp <- slope_tidy %>% 
  dplyr::group_by(id) %>% 
  dplyr::mutate(
    MO2_raw = abs(mo2) * net_volume * 60 * 60,
    MO2_bground_corr = abs(mo2 - bground) * net_volume * 60 * 60,
    MO2_corr = abs(mo2corr) * net_volume * 60 * 60
  ) %>% 
  dplyr::select(id, order, DO, MO2_raw, MO2_bground_corr, MO2_corr) %>% 
  tidyr::pivot_longer(
    cols = c(MO2_raw, MO2_bground_corr, MO2_corr), 
    names_to = "MO2_type"
  ) %>%
  dplyr::mutate(
    MO2_type = forcats::fct_relevel(MO2_type, "MO2_raw", "MO2_bground_corr", "MO2_corr")
  ) %>%
  dplyr::ungroup()
```

### Figure S1

Now let's plot them in the same graph for each fish: 
<span style="color:#f692B4">raw MO₂ calculations</span> are shown in pink, 
<span style="color:#6E7889">MO₂ calculations after correcting for background respiration</span> are shown in grey, and 
<span style="color:#53af8b">MO₂ calculations after correcting for background respiration and oxygen diffusion dynamics</span> are shown in green.


```{r, fig.height = 8, fig.width= 10}
# Define your custom colour scheme
plot_colours <- c("#f692B4", "#6E7889", "#53af8b")

# Collect unique fish IDs
fish_ids <- slope_corr_comp %>% 
  dplyr::distinct(id) %>% 
  dplyr::pull(id)

# Create a list to store plots
plot_list <- list()

# Generate plots and add them to the list
for (id_i in fish_ids) {
  plot_i <- slope_corr_comp %>% 
    dplyr::filter(id == id_i) %>% 
    ggplot(aes(x = DO, y = value, colour = MO2_type)) +
    geom_point(alpha = 0.5) +
    theme_clean() +
    labs(
      subtitle = id_i,
      x = "Dissolved oxygen percentage (DO)",
      y = "MO2 (O2 mg/g/h)"
    ) +
    scale_colour_manual(values = plot_colours) +
    theme(legend.position = "none")
  
  plot_list[[id_i]] <- plot_i
}

# Combine plots in pages of 9 (3 columns x 3 rows)
plots_per_page <- 9
n_pages <- ceiling(length(plot_list) / plots_per_page)

# Display each page
for (i in seq_len(n_pages)) {
  start <- (i - 1) * plots_per_page + 1
  end <- min(i * plots_per_page, length(plot_list))
  page_plots <- patchwork::wrap_plots(plot_list[start:end], ncol = 3)
  print(page_plots)
}
```


**Figure S1**: Individual fish-level plots showing the relationship between dissolved oxygen (DO) percentage and metabolic oxygen consumption rate (MO₂, in mg O₂/g/h) for each 20-min interval. Each panel displays data for a single fish (fish ID in the top left of the plot). Points are colour-coded by the type of MO₂ calculation:
<span class="pink">raw MO₂</span> (pink, #f692B4),
<span class="grey">background-corrected MO₂</span> (grey, #6E7889), and
<span class="green">fully corrected MO₂ (accounting for background respiration and oxygen diffusion dynamics)</span> (green, #53af8b).
Each point represents a measurement at a specific DO level, and the slight transparency (alpha = 0.5) highlights overlapping observations.

<!------------------------------->

# 📊 Exploratory Data Analysis (EDA)

<!------------------------------->

## O₂ vs MO₂

### Figure S2

This interactive was used to identify any outliers, or potential errors. This has not been generated in the PDF `Supplementary file 1`, as it is not compatible. 

```{r}
lm_lines <- slope_tidy %>%
  group_by(id) %>%
  summarise(
    DO_seq = list(seq(min(DO), max(DO), length.out = 100)),  # Generate a sequence of DO values
    MO2_pred = list(predict(lm(MO2_g ~ DO, data = cur_data()), newdata = data.frame(DO = seq(min(DO), max(DO), length.out = 100))))
  ) %>%
  unnest(c(DO_seq, MO2_pred))  # Expand lists into rows

# Create scatter plot with markers for each fish
p <- plot_ly(
  data = slope_tidy,
  x = ~DO,
  y = ~MO2_g,
  type = "scatter",
  mode = "markers",
  color = ~id,  # Colour points by fish ID
  marker = list(opacity = 0.6),
  name = ~id
)

# Add regression lines for each fish
p <- p %>%
  add_trace(
    data = lm_lines,
    x = ~DO_seq,
    y = ~MO2_pred,
    type = "scatter",
    mode = "lines",
    color = ~id,  # Ensure each line matches its corresponding fish
    line = list(width = 1, dash = "solid"),
    showlegend = FALSE  # Avoid cluttering the legend
  )

# Final layout
p <- p %>%
  layout(
    title = "MO<sub>2</sub> vs Dissolved Oxygen with individual linear regressions",
    xaxis = list(title = "Dissolved Oxygen (%)"),
    yaxis = list(title = "MO<sub>2</sub> (mg<sup>-1</sup> O<sub>2</sub> g<sup>-1</sup> h<sup>-1</sup>)"),
    showlegend = FALSE
  )

# Display plot
p
```

***Figure S2:*** interactive plot of metabolic rate measurements (MO₂; mg O₂ g^-1^h^-1^) by dissolved oxygen percentage (DO) for all fish, including all estimates during the SMR phase (i.e. intermittent phase). Individual linear regression were fitted for visual reference, and do not represent the best fitting regression.

### Figure S3

Looking at the difference responses in the two salinity groups.

```{r}
salinity_summary <- slope_tidy %>% 
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe(n = length(unique(id)))

slope_tidy %>% 
  ggplot(aes(y = MO2_g, x = DO, colour = id)) + 
  geom_point(show.legend = FALSE) +
  geom_smooth(aes(group = id), method = "lm", se = FALSE, colour = scales::alpha("black", 0.5)) + # Transparent black lines
  geom_smooth(method = "lm", se = TRUE, colour = "red") + # Overall smooth line
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), se = TRUE, colour = "red", linetype = "dashed") +
  theme_clean() +
  facet_wrap(~salinity_group) +
  labs(
    subtitle = "mo2 vs o2 by salinity treatment",
    x = "Dissolved oxygen percentage (DO)",
    y = "MO2 (O2 mg/g/h)"
  ) +
  geom_text(data = salinity_summary,
            aes(x = -Inf, y = Inf, label = paste0("italic(n) == ", n)),
            hjust = -0.1, vjust = 1.2, inherit.aes = FALSE, parse = TRUE)
```

***Figure S3:*** Metabolic rate measurements (MO₂; mg O₂ g^-1^h^-1^) by dissolved oxygen percentage (DO) for fish from the two salinity treatments, including all estimates during the SMR phase (i.e. intermittent phase). Individual linear regression were fitted for visual reference, and do not represent the best fitting regression. The solid red and dashed red line represents the global trend in each of the treatments, both a linear (solid red) and non-linear (dashed red; Generalized Additive Model fitted with `geom_smooth`).

### Figure S4

A plot to look at the different chamber types

```{r}
chamber_summary <- slope_tidy %>% 
  dplyr::group_by(chamber_type) %>% 
  dplyr::reframe(n = length(unique(id)))

slope_tidy %>% 
  ggplot(aes(y = MO2_g, x = DO, colour = id)) + 
  geom_point(show.legend = FALSE) +
  geom_smooth(aes(group = id), method = "lm", se = FALSE, colour = scales::alpha("black", 0.5)) + # Transparent black lines
  geom_smooth(method = "lm", se = TRUE, colour = "red") + # Overall smooth line
  geom_smooth(se = TRUE, colour = "red", linetype = "dashed") +
  theme_clean() +
  facet_wrap(~chamber_type, scale = "free") +
  labs(
    subtitle = "mo2 vs o2 by chamber type",
    x = "Dissolved oxygen percentage (DO)",
    y = "MO2 (mg O2 g/h)"
  ) +
  geom_text(data = chamber_summary,
            aes(x = -Inf, y = Inf, label = paste0("italic(n) == ", n)),
            hjust = -0.1, vjust = 1.2, inherit.aes = FALSE, parse = TRUE)
```

***Figure S4:*** Metabolic rate measurements (MO₂; mg O₂ g^-1^h^-1^) by dissolved oxygen percentage (DO) for fish tested in the 4 different chamber types, including all estimates during the SMR phase (i.e. intermittent phase). Individual linear regression were fitted for visual reference, and do not represent the best fitting regression. The solid red and dashed red line represents the global trend in each of the treatments, both a linear (solid red) and non-linear (dashed red; Generalized Additive Model fitted with `geom_smooth`).


## Routine MO₂

Making an SMR phase only data frame

```{r}
slope_tidy_smr <- slope_tidy %>% 
  dplyr::filter(phase == "smr")
```


Routine MO<sub>2</sub> by salinity. Plot of all MO<sub>2</sub> measures during SMR phase by salinity treatment. The small points are the raw observed values, the shaded area behind the points is the a kernel density of the observed data, the box plot shows the median and interquartile range (IQR), and the large point shows the mean.

```{r}
mean_mo2_salinity <- slope_tidy_smr %>% 
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe(mean_mo2 = mean(MO2, na.rm = TRUE))

fig_i <- ggplot() +
    geom_violin(data = slope_tidy_smr, aes(x = salinity_group, y = MO2, fill = salinity_group), color = NA, alpha = 0.3) +
  geom_jitter(data = slope_tidy_smr, aes(x = salinity_group, y = MO2, fill = salinity_group),
                       shape = 21, size = 2, color = "black", alpha = 0.2) +
  geom_boxplot(data = slope_tidy_smr, aes(x = salinity_group, y = MO2, fill = salinity_group),
                        size = 1, alpha = 0.5, outlier.shape = NA, width = 0.3) +
  geom_point(data = mean_mo2_salinity, 
                aes(x = salinity_group, y = mean_mo2, fill = salinity_group), 
                size = 3, alpha = 0.8, colour = "black", stroke = 2) +
  scale_fill_manual(values = c("#4B5320", "#000080")) +  # Custom fill colours
  scale_colour_manual(values = c("#4B5320", "#000080")) +
  theme_clean() +
  theme(legend.position = "none") +
  labs(
    subtitle = "",
    x = "Salinity group (ppt)",
    y = "Routine MO2 (mg O2 g/h)"
  )

fig_i
```


## SMR

Here's the same plot but for only the SMR, as estimated with calcSMR function by Chabot, Steffensen and Farrell (2016)^[1]^. Specifically, we use mean of the lowest normal distribution (MLND) where CVmlnd \< 5.4 and the mean of the lower 20% quantile (q0.2) were CVmlnd \> 5.4. If CVmlnd is not calculated we have used q0.2.

```{r}
smr_only <- slope_tidy_smr %>% 
  dplyr::group_by(id) %>% 
  dplyr::slice(1)

mean_smr_salinity <- smr_only %>% 
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe(mean_smr = mean(SMR, na.rm = TRUE))

fig_i <- ggplot() +
    geom_violin(data = smr_only, aes(x = salinity_group, y = SMR, fill = salinity_group), color = NA, alpha = 0.3) +
  geom_jitter(data = smr_only, aes(x = salinity_group, y = SMR, fill = salinity_group),
                       shape = 21, size = 2, color = "black", alpha = 0.2) +
  geom_boxplot(data = smr_only, aes(x = salinity_group, y = SMR, fill = salinity_group),
                        size = 1, alpha = 0.5, outlier.shape = NA, width = 0.3) +
  geom_point(data = mean_smr_salinity, 
                aes(x = salinity_group, y = mean_smr, fill = salinity_group), 
                size = 3, alpha = 0.8, colour = "black", stroke = 2) +
  scale_fill_manual(values = c("#4B5320", "#000080")) +  # Custom fill colours
  scale_colour_manual(values = c("#4B5320", "#000080")) +
  theme_clean() +
  theme(legend.position = "none") +
  labs(
    subtitle = "",
    x = "Salinity group (ppt)",
    y = "MO2 (mg O2 g/h)"
  )

fig_i
```

## Individual O₂, MO₂, and SMR

Here we will plot the individual relationship between O₂, MO₂, and SMR for all fish

Create output directory if needed

```{r, echo=FALSE}
# Create output directory if needed
output_fig_slopes_wd <- file.path(output_fig_wd, "slopes")
if (!dir.exists(output_fig_slopes_wd)) {
  dir.create(output_fig_slopes_wd)
}
```

Loop through each fish ID to create a plot, save these to a single PDF file, this is called `combined_slopes.pdf`. It can be found in this path `output-fig/slopes/combined_slopes.pdf` 
These plots show Metabolic rate measurements (MO₂; mg O₂ g^-1^h^-1^ by O<sub>2</sub>) at each of the four experimental phase (the over night SMR intermittent-flow respirometry phase, closed phases at 50%, 75% or 100% O<sub>2</sub>). The estimated SMR value is represented as a dashed red line.

```{r, fig.height = 3, fig.width= 4}
# Define consistent colors for all phases
phase_colors <- c(
  "50c"  = "#e56b6f",
  "75c"  = "#b56576",
  "100c" = "#6d597a",
  "smr"  = "#355070"
)

ids <- slope_tidy %>%
  dplyr::distinct(id) %>%
  pull(id) %>%
  as.list()

MO2_plot_list <- list()

# 1) Open the PDF device once
pdf(
  file   = file.path(output_fig_slopes_wd, "combined_slopes.pdf"),
  width  = 8,
  height = 6
)

# 2) Loop over IDs and create each plot
for (id_i in ids) {

  smr <- slope_tidy %>%
    dplyr::filter(id == id_i) %>%
    dplyr::slice(1) %>%
    dplyr::pull(SMR)

  plot <- slope_tidy %>%
    dplyr::filter(id == id_i) %>%
    ggplot(aes(x = o2, y = MO2)) +
    geom_hline(yintercept = smr, linetype = "dashed", color = "darkred") +
    geom_point(aes(colour = phase)) +
    scale_color_manual(values = phase_colors, drop = FALSE) +  # Ensures consistent colours
    theme_clean() +
    labs(
      subtitle = paste0(id_i, " slopes"),
      x = "Mean o2 (mg_per_l)",
      y = "abs(mo2) (mg_per_l)"
    )

  # Instead of saving each plot separately, just print it
  print(plot)

  MO2_plot_list[[id_i]] <- plot
}

# 3) Close the PDF device *after* the loop
dev.off()
```

<!------------------------------->

# 🧮 Analysis

<!------------------------------->

## Routine MO2

### Scaling predictors

Here we scale our predictors for the model. We have first applied an log transformation to mass, so its directly interpretable as the mass-scaling exponent, just like in metabolic theory.

```{r}
scale_list <- c("temp", "order", "log_mass") 

slope_tidy_smr <- slope_tidy_smr %>%
  dplyr::mutate(log_mass = log(mass),
                across(all_of(scale_list), ~ scale(.x, center = TRUE, scale = FALSE),
                       .names = "{.col}_c"),
                light_dark_c = if_else(light_dark == "light", 0.5, -0.5))
```

Summary of predictors used in the model below

```{r}
slope_tidy_smr %>%
  dplyr::mutate(cycle = as.numeric(cycle)) %>% 
  dplyr::reframe(temp_range = paste0(round(min(temp, na.rm = TRUE),2), "–", round(max(temp, na.rm = TRUE),2)),
            mass_range = paste0(round(min(mass, na.rm = TRUE),2), "–", round(max(mass, na.rm = TRUE),2)),
          cycle_range = paste0(round(min(cycle, na.rm = TRUE),2), "–", round(max(cycle, na.rm = TRUE),2)),
          number_fish = length(unique(id))  
          ) %>% 
  gt()
```

### Model structure

Here we will use a Bayesian Generalised Linear Mixed Model (GLMM) with a Gamma distribution and a log link, where the shape parameter (α) is also modelled as a function of predictors. This models MO<sub>2</sub> by salinity during the SMR phase to see if the fish held at different salinities have different RMRs. We have also added a few scaled predictors, that may help describe variation in the data, such as mass (g; 0.21--1.6) temperature (°C; 13.84--14.38), cycle order (5--27), and light/dark cycle (light or dark; light between 07:00:00 and 19:00:00), we also include a random effect for fish id to account for multiple MO<sub>2</sub> measures on each fish (1--58). We allowed the the shape parameter (α) to vary as a function of some of the predictors (e.g. salinity_group, order_z) to improve fit.

```{r}
mo2_gamma_bf <- bf(MO2 ~ temp_c + 
                     order_c  + 
                     light_dark_c + 
                     log_mass_c + 
                     salinity_group + (1|id),
                   shape ~ salinity_group + 
                     order_c,
                   family = Gamma(link = "log"))
```

### Prior selection

These are the default priors. We will use these.

```{r}
default_prior <- get_prior(mo2_gamma_bf, data = slope_tidy_smr, family = Gamma(link = "log"))
default_prior %>% 
  gt()
```

### Run model

Here we run the model, I have hashed this out because I have saved the model for quick reloading.

⏭️ **Skip this step** if you have downloaded the saved model `mo2_mod_gamma.rds` from GitHub (or pulled the entire GitHub project)

```{r, echo=FALSE}
# mo2_mod_gamma <- brm(mo2_gamma_bf,
#                data = slope_tidy_smr,
#                cores = 4,
#                chains = 4,
#                prior = default_prior,
#                warmup = 1000,
#                seed = 143019,
#                thin = 2,
#                iter = 8000,
#                save_pars = save_pars(all=TRUE),
#                sample_prior = TRUE,
#                file = paste0(output_mods_wd, './mo2_mod_gamma'))
# print("Model complete")
```

Here we reload the model

```{r}
mo2_mod_gamma <-  readRDS(file = paste0(output_mods_wd, './mo2_mod_gamma.rds'))
```

### Model diagnostics

Checking model convergence

```{r}
color_scheme_set("red")
plot(mo2_mod_gamma, ask = F)
```

Checking rhat are equal to one

```{r}
summary(mo2_mod_gamma)
```

Using leave one out (loo) measure of fit, the model appears to preform well, all, but one Pareto k estimates are good (k \< 0.7)

```{r}
loo(mo2_mod_gamma)
```

Model predictions generally align with the observed data

```{r}
color_scheme_set("red")
p <- pp_check(mo2_mod_gamma, type = "dens_overlay")
p
```

### 📈 Results

We did not see a meaningful difference between the routine metabolic rate for fish from the two salinity treatments.

#### Table S1

**Table S1**: Fixed effect Estimates (β) and 95% Credible Intervals (95% CI)

```{r}
model_est <- fixef(mo2_mod_gamma, probs = c(0.025, 0.975)) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "Predictor") %>% 
  dplyr::mutate('β' = round(Estimate, 3),
                Q2.5 = round(Q2.5, 3),
                Q97.5 = round(Q97.5, 3),
                '95% CI' = paste0("[", Q2.5, ", ", Q97.5, "]"))

model_est %>% 
  dplyr::select(Predictor, 'β', '95% CI') %>% 
  gt()
```

Looking at the marginal mean MO2 for each salinity treatment

```{r}
em_results <- emmeans(mo2_mod_gamma, ~ salinity_group) 

em_results_df <- em_results %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ round(exp(.),3)))

em_results_df %>% 
  gt()
```

These are the estimated mean constrast, we have exponentiated the log estimate to get the fold-change. This means the model predicts that sailianty group 0 has \~12% higher metabolic rate than group 9 (1.12)---but this difference is not credible, because the interval for that difference includes no difference (i.e. a ratio of 1.0).

```{r}
contrast_results <- contrast(em_results, method = "pairwise")

contrast_results_df <-  contrast_results %>% 
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ round(exp(.),2)))

contrast_results_df %>% 
  gt()
```

EEM contrasts for the two light phases, where -0.5 is dark, and 0.5 is light

```{r}
em_results <- emmeans(mo2_mod_gamma, ~ light_dark_c,
                       at = list(light_dark_c = c(0.5, -0.5)))

contrast_results <- contrast(em_results, method = "pairwise", .rev = TRUE)

contrast_results_df <-  contrast_results %>% 
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ round(exp(.),2)))

contrast_results_df %>% 
  gt()
```

Here we we calculate the model-estimated marginal means for MO2 for a unit increase in mass and transform this onto a fold-change scale. Where log(mass) has be centered, a vaule of 0 represents the mean mass of fish, and 1 corresponds to a fish with log(mass) = 1 unit above the mean (i.e. 2.7 times large).

log(mass<sub>2</sub>) = log(mass<sub>1</sub>)+1

If you exponentiate both sides:

mass<sub>2</sub> = mass<sub>1</sub> x exp(1)

mass<sub>2</sub> ≈ mass<sub>1</sub> x 2.718

```{r}
em_mass <- emmeans(mo2_mod_gamma, ~ log_mass_c, at = list(log_mass_c = c(0, 1)))

# Get the contrast (difference in log MO₂), then back-transform
contrast(em_mass, method = "revpairwise") %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), exp)) %>%
  mutate(across(where(is.numeric), ~ round(., 2))) %>%
  gt()
```

Pulling the emmeans draws for our plot

```{r}
emmeans_draws_rmr <- mo2_mod_gamma %>%
  emmeans(~ salinity_group) %>%
  gather_emmeans_draws() %>% 
  dplyr::mutate(.value =  exp(.value),
                salinity_group = as.character(salinity_group))

emmeans_contrast_draws_rmr  <-  mo2_mod_gamma %>%
  emmeans(~ salinity_group) %>%
  contrast(method = "pairwise") %>%
  gather_emmeans_draws() %>% 
  dplyr::mutate(.value =  exp(.value))
```

#### Figure 1a

***NOTE: This plot is in the main text of the manuscript, which is way it is called Figure 1a (not Figure Si)***

```{r}
fig_1a <- ggplot() +
  geom_violin(data = slope_tidy_smr,
              aes(x = salinity_group, y = MO2, fill = salinity_group),
              color = NA, alpha = 0.2) +
  geom_jitter(data = slope_tidy_smr,
              aes(x = salinity_group, y = MO2, fill = salinity_group),
              shape = 21, width = 0.3, size = 1, color = "black", alpha = 0.1) +
  geom_point(data = mean_mo2_salinity,
             aes(x = salinity_group, y = mean_mo2, fill = salinity_group),
             size = 4, alpha = 1, stroke = 2, color = "black", shape = 21,
             position = position_nudge(x = 0.05)) +
  stat_pointinterval(data = emmeans_draws_rmr, 
                     aes(x = salinity_group, y = .value),
                     color = "black", fill = "grey", point_interval = "mean_qi", .width = 0.95, shape = 21,  stroke = 2, point_size = 4, alpha = 1,
                     position = position_nudge(x = -0.05)) +
  scale_y_continuous(limits = c(0.00, 0.35)) +  # 👈 Set y-axis range here
  scale_fill_manual(values = c("#4B5320", "#000080")) +
  scale_colour_manual(values = c("#4B5320", "#000080")) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA)
  ) +
  labs(
    subtitle = "",
    x = "Salinity group (ppt)",
    y = "Routine MO2 (mg O2 g/h)"
  )

fig_1a
```

**Figure 1a:** Routine metabolic rate (i.e. MO<sub>2</sub> (mg^-1^ O<sub>2</sub> h^-1^) measured during SMR meassurments) plotted by salinity treatment. The small transparent points are the observed values, the shaded area behind the points is the a kernel density of the observed data, the large coloured point (to the right) is the observed mean, the large grey point with error bars (to the left) is the model estimated marginal mean (eemean) 95% Credible Intervals (95% CI).

```{r}
# ggsave(filename = paste0(output_fig_wd, "./ms_figures./figure-1a.pdf"),
#        plot = fig_1a,
#        width = 210/2,
#        height = 297/2,  # Specify the height of the plot
#        units = "mm")
```


There was moderate but meaningful variation in RMR between individuals, even after accounting for mass, temperature, salinity, and light phase. The estimated standard deviation of individual RMR was 0.31 [0.26 to 0.38], indicating that individual identity explained a meaningful proportion of the observed variation.

```{r}
random_effect <- summary(mo2_mod_gamma)$random$id
random_effect %>% 
  gt()
```

## SMR

### Formating and scaling data

Here we are filtering the data frame to have only measure per fish for the SMR estimate

```{r}
scale_list <- c("temp_mean", "log_mass", "cycles")

smr <- slope_tidy_smr %>%
  dplyr::group_by(id) %>% 
  dplyr::reframe(temp_mean = mean(temp),
                 log_mass = log_mass[1],
                 mass = mass[1],
                 SMR = SMR[1],
                 salinity_group = salinity_group[1],
                 cycles = length(order)) %>% 
  dplyr::mutate(across(all_of(scale_list), ~ scale(.x, center = TRUE, scale = FALSE),
                       .names = "{.col}_c"))
```

Summary of predictors used in the model below

```{r}
smr %>%
  dplyr::reframe(temp_range = paste0(round(min(temp_mean, na.rm = TRUE),2), "–", round(max(temp_mean, na.rm = TRUE),2)),
            mass_range = paste0(round(min(mass, na.rm = TRUE),2), "–", round(max(mass, na.rm = TRUE),2)),
          cycle_range = paste0(round(min(cycles, na.rm = TRUE),2), "–", round(max(cycles, na.rm = TRUE),2)),
          number_fish = length(unique(id))  
          ) %>% 
  gt()
```

### Model structure

Here we will use a Bayesian Generalised Linear Mixed Model (GLMM) with a Gamma distribution and a log link `Gamma(link = "log")`, where the shape parameter (α) is also modelled as a function of the salinity group, `shape ~ salinity_group`. This models MO2 by salinity during the SMR phase to see if the fish held at different salinities have different SMRs. We have also added a few scaled predictors, that may help describe variation in the data, such as fish mass (g; 0.21--1.6) mean temperature (°C; 13.841--14.277), and the number of cycles over which SMR was estimated (1--23). We allowed the the shape parameter (α) to vary as a function of salinity_group to improve fit.

```{r}
smr_gamma_bf <- bf(SMR ~ temp_mean_c + 
                  cycles_c  +
                  log_mass_c + 
                  salinity_group,
                  shape ~ salinity_group,
                  family = Gamma(link = "log"))
```

### Prior selection

These are the default priors for the model. We will use these.

```{r}
priors_default <- get_prior(smr_gamma_bf, data = smr, family = Gamma(link = "log"))
priors_default %>% 
  gt()
```

### Run model

Here we run the model, I have hashed this out because I have saved the model for quick reloading.

⏭️ **Skip this step** if you have downloaded the saved model 'smr_mod_gamma.rds'

```{r, echo=FALSE}
# smr_mod_gamma_c <- brm(smr_gamma_bf,
#                data = smr,
#                cores = 4,
#                chains = 4,
#                prior = priors_default,
#                warmup = 1000,
#                seed = 143019,
#                thin = 2,
#                iter = 8000,
#                save_pars = save_pars(all=TRUE),
#                sample_prior = TRUE,
#                file = paste0(output_mods_wd, './smr_mod_gamma'))
# print("Model complete")
```

Here we reload the model

```{r}
smr_mod_gamma <-  readRDS(file = paste0(output_mods_wd, './smr_mod_gamma.rds'))
```

### Model diagnostics

Checking model convergence

```{r}
color_scheme_set("red")
plot(smr_mod_gamma, ask = F)
```

Using leave one out (loo) measure of fit, the model appears to preform well, two Pareto k estimates falls outside the good range (0.7, 1]

```{r}
loo(smr_mod_gamma)
```

Model predictions generally align with the observed data, but there is a lot of uncertainty around this estimate.

```{r}
color_scheme_set("red")
p <- pp_check(smr_mod_gamma, type = "dens_overlay")
p
```

### 📈 Results

We did not see a meaningful difference between the routine metabolic rate for fish from the two salinity treatments.

#### Table S2

**Table S2:** Fixed effect Estimates (β) and 95% Credible Intervals (95% CI) from a Bayesian Generalised Linear Mixed Model (GLMM) with a Gamma distribution and a log link

```{r}
model_est <- fixef(smr_mod_gamma, probs = c(0.025, 0.975)) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "Predictor") %>% 
  dplyr::mutate('β' = round(Estimate, 3),
                Q2.5 = round(Q2.5, 3),
                Q97.5 = round(Q97.5, 3),
                '95% CI' = paste0("[", Q2.5, ", ", Q97.5, "]"))

model_est %>% 
  dplyr::select(Predictor, 'β', '95% CI') %>% 
  gt()
```

Looking at the marginal mean difference between salinity groups

```{r}
em_results <- emmeans(smr_mod_gamma, ~ salinity_group) 

em_results_df <-  em_results %>% 
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ round(exp(.),3)))


em_results_df %>% 
  gt()
```

```{r}
contrast_results_df <-  contrast_results %>% 
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ exp(.)))

contrast(em_results, method = "pairwise") %>% 
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~ round(exp(.),2))) %>% 
  gt()
```

Here we we calculate the model-estimated marginal means for SMR for a unit increase in mass and transform this onto a fold-change scale. Where log(mass) has be centered, a vaule of 0 represents the mean mass of fish, and 1 corresponds to a fish with log(mass) = 1 unit above the mean (i.e. 2.7 times large).

```{r}
em_mass <- emmeans(smr_mod_gamma, ~ log_mass_c, at = list(log_mass_c = c(0, 1)))

# Get the contrast (difference in log MO₂), then back-transform
contrast(em_mass, method = "revpairwise") %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), exp)) %>%
  mutate(across(where(is.numeric), ~ round(., 2))) %>%
  gt()
```

Pulling the emmeans draws for our plot

```{r}
emmeans_draw_smr <- smr_mod_gamma %>%
  emmeans(~ salinity_group) %>%
  gather_emmeans_draws() %>% 
  dplyr::mutate(.value =  exp(.value),
                salinity_group = as.character(salinity_group))

emmeans_contrast_draws_smr <-  smr_mod_gamma %>%
  emmeans(~ salinity_group) %>%
  contrast(method = "pairwise") %>%
  gather_emmeans_draws() %>% 
  dplyr::mutate(.value =  exp(.value))
```

#### Figure 1b

**NOTE: This plot is in the main text of the manuscript which is why it is Figure 1b (not Figure Si)**

```{r}
mean_smr_salinity <- smr %>%
  dplyr::group_by(salinity_group) %>%
  dplyr::reframe(mean_SMR = mean(SMR, na.rm = TRUE))


fig_1b <- ggplot() +
  geom_violin(data = smr,
              aes(x = salinity_group, y = SMR, fill = salinity_group),
              color = NA, alpha = 0.2) +
  geom_jitter(data = smr,
              aes(x = salinity_group, y = SMR, fill = salinity_group),
              shape = 21, width = 0.3, size = 1, color = "black", alpha = 0.1) +
  geom_point(data = mean_smr_salinity,
             aes(x = salinity_group, y = mean_SMR, fill = salinity_group),
             size = 4, alpha = 1, stroke = 2, color = "black", shape = 21,
             position = position_nudge(x = 0.05)) +
  stat_pointinterval(data = emmeans_draw_smr,
                     aes(x = salinity_group, y = .value),
                     color = "black", fill = "grey", point_interval = "mean_qi", .width = 0.95, shape = 21,  stroke = 2, point_size = 4, alpha = 1,
                     position = position_nudge(x = -0.05)) +
  scale_y_continuous(limits = c(0.00, 0.35)) +
  scale_fill_manual(values = c("#4B5320", "#000080")) +  # Custom fill colours
  scale_colour_manual(values = c("#4B5320", "#000080")) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA)
  ) +
  labs(
    subtitle = "",
    x = "Salinity group (ppt)",
    y = "Standard metabolic rate (SMR; mg O2 g/h)"
  )

fig_1b
```


**Figure 1b:** The standard metabolic rate estimate (mg^-1^ O<sub>2</sub> h^-1^) plotted by salinity treatment. The small transparent points are the observed values, the shaded area behind the points is the a kernel density of the observed data, the large coloured point (to the right) is the observed mean, the large grey point with error bars (to the left) is the model estimated marginal mean (eemean) 95% Credible Intervals (95% CI).

```{r}
# ggsave(filename = paste0(output_fig_wd, "./ms_figures./figure-1b.pdf"),
#        plot = fig_1b,
#        width = 210/2,
#        height = 297/2,  # Specify the height of the plot
#        units = "mm")
```

## Incremental regression analyses

Here we are following the methods Urbina et al. (2012)^[1]^ with an incremental regression analyses, in order to determine the best fit for the MO<sub>2</sub> vs O<sub>2</sub> data

This analysis approach evaluates the relative 'fit' of each polynomial order equation starting at zero and increasing to the third order, permitting a mathematical assessment of whether fish were oxyconforming or oxyregulatoring. If the data is best fitted/predicted by a single linear relationship (1^st^-order polynomial) with a positive slope, this would suggests the fish were oxyconforming. Alternately, if the relationship is best modelled by a flat regression (0^th^-order polynomial), or a higher order polynomial (2^nd^ or 3^rd^-order polynomial) the fish is likely oxyregulatoring.

### Building Bayesian regressions

Here we are using a Bayesian approach to model fitting with brm. These models take a long time to run, so I have saved them and re-loaded them to save time. I have also saved the summary data produced from the models, to save time, you can simply skip the hashed code and input the resulting summary data.

We will run our custom function, `bayes_incremental_regression_by_id`.

Let's make define the output directory

```{r}
output_mods_bayes_wd <- paste0(wd, "./output-mod./bayes-regs")
```

⏭️ **Skip this step** if you have already run this once, or have downloaded the saved models or saved data files from GitHub (that's why its hashed out). ⌛*This code takes a while to run *⌛

During this part of the script I received an error, 'Caused by error in `socketConnection()`:'. I think the system may be hitting a limit on the number of simultaneous socket connections. If you this part of the code and also get this issue, try reducing the number of parallel workers: plan(multisession, workers = 2), or run again after the error.

```{r, results='hide'}
# ids <- slope_tidy %>%
#   dplyr::distinct(id) %>%
#   pull(id)
# 
# plan(multisession)
# 
# future_map(
#   ids,
#   bayes_incremental_regression_by_id,
#   id_name = "id",
#   data = slope_tidy,
#   response = "MO2_g",
#   predictor = "DO",
#   seed_number = 143019,
#   save_models = TRUE,
#   mod_output_wd = output_mods_bayes_wd
# )
# 
# plan(sequential)
```

Load all models and store in a list, will use a lot of memory. You can also skip this step and load the resulting data frames below. I am using the custom function `load_rds`, so we can compare them and generate predictions.

⏭️ **Skip this step** if you have downloaded the saved data pulled from these models have also hased this out, bayes_reg_mods_fit.csv' and 'bayes_reg_mods_predictions.csv'. These are in the mod-data directory.

```{r}
# bayes_reg_mods <- load_rds(model_dw = output_mods_bayes_wd)
```

### Model fits

⏭️ **Skip this step** if you have downloaded **💿 bayes_reg_mods_fit.csv**.

It gets model fit parameters loo and r2 using the custom function, `incremental_regression_bayes_fits`. *⌛ This code takes a while to run ⌛*

```{r}
# bayes_reg_mods_fit <- incremental_regression_bayes_fits(models = bayes_reg_mods)
# write.csv(bayes_reg_mods_fit, paste0(mod_data_wd, "./bayes_reg_mods_fit.csv"), row.names = FALSE)
```

Reading in this model fit data frame, in the case you did not load in all the models.

```{r}
bayes_reg_mods_fit <- read.csv(paste0(mod_data_wd, "./bayes_reg_mods_fit.csv"))
```

**Selecting the best fitting model**

elpd_loo, or the expected log pointwise predictive density for leave-one-out cross-validation, is a metric used in Bayesian model evaluation to assess the predictive accuracy of a model. The elpd_loo is an approximation of how well the model is expected to predict new data, based on leave-one-out cross-validation. Higher elpd_loo values indicate better predictive performance.

```{r}
best_fit_bayes_reg <- bayes_reg_mods_fit %>%
  dplyr::group_by(id) %>%
  dplyr::mutate(elpd_loo_rank = rank(-elpd_loo)) %>%
  dplyr::select(id, model_type, elpd_loo, r2, elpd_loo_rank, r2_q2.5, r2_q97.5, estimate_DO, conf.low_DO, conf.high_DO) %>%
  ungroup()
```

### Model predictions

⏭️ **Skip this step** if you have downloaded **💿 bayes_reg_mods_predictions.csv**.

It pulls our model predictions using a custom function `bayes_mod_predictions`.

```{r}
# bayes_reg_mods_predictions <- bayes_mod_predictions(models = bayes_reg_mods, original_data = slope_tidy)
# write.csv(bayes_reg_mods_predictions, paste0(mod_data_wd, "./bayes_reg_mods_predictions.csv"), row.names = FALSE)
```

Reading in the predicted data

```{r}
bayes_reg_mods_predictions <- read.csv(paste0(mod_data_wd, "./bayes_reg_mods_predictions.csv"))
```

We are going to combined this with our best fitting model df, so we know how they ranks for LOO.

```{r}
bayes_reg_mods_predictions <- left_join(bayes_reg_mods_predictions, best_fit_bayes_reg, by = c("id", "model_type"))
```

### 📈 Results

#### Model selection summary

In most cases, the relationship between MO<sub>2</sub> and DO was best modelled with a 2^nd^-order polynomial (*n* = 22, 38% of fish), followed by a 3^rd^-order polynomial (*n* = 15, 26%), 1^st^-order polynomial (*n* = 11, 19%), and finally 0^th^-order polynomial (*n* = 10, 17%).

For the two most common models, 2^nd^- and 3^rd^-order polynomials, this could suggest the presence of a critical oxygen threshold (Pcrit) where the relationship between MO<sub>2</sub> and O<sub>2</sub> changes. To confirm the presence of a Pcrit, we need to validated the shape of the polynomials (see Pcrit model below). In any case, these type of models are indicative of **oxyregulator**.

The next most common model was is 1^st^-order polynomial. In the case of the 1^st^-order polynomials, it suggest the presences of linear relationship between o2 and MO2, which is indicative of **oxyconformer**. However, to be true evidence of a oxyconformer this relationship should be positive (i.e. as O<sub>2</sub> falls MO<sub>2</sub> also falls). 13 of the 18 individuals best modelled with a linear function had positive estimates with credible intervals that did not overlap with zero (Table S3).

Lastly, 0^th^-order was the least common (*n* = 3, 5%), it suggests that MO<sub>2</sub> does not show a statistically significant dependence on the O<sub>2</sub>. In other words, the metabolic rate does not adjust based on oxygen availability, and there is no clear critical oxygen threshold (Pcrit) where the relationship changes. This is indicative of a **oxyregulator**.

```{r}
best_mod <- best_fit_bayes_reg %>% 
  dplyr::filter(elpd_loo_rank == 1)

total_fish <- nrow(best_mod)

mod_summary_table <- best_mod %>% 
  dplyr::group_by(model_type) %>% 
  dplyr::reframe(n = length(id),
                 percent = round((n/total_fish)*100,2)) %>% 
  dplyr::mutate(best_model_name = case_when(
      model_type == "lm_0" ~ "0th-order polynomial",
      model_type == "lm_1" ~ "1st-order polynomial",
      model_type == "lm_2" ~ "2nd-order polynomial",
      model_type == "lm_3" ~ "3rd-order polynomial",
      TRUE ~ "ERROR"
    ))


table_bwm <- mod_summary_table %>% 
  dplyr::select(best_model_name, everything(), -model_type) %>% 
   gt() %>% 
  cols_align(
    align = "center", 
    columns = everything()
  )

table_bwm
```

Summary of fish best model with a linear function.

### Table S3

**Table S3:** Ten fish that were best modelled with a linear function, showing r^2^, and estimate (). Only two fish had positive estimates with credible intervals that did not overlap with zero, which are highlighted as conforming in the table. Thus in total, 13 of 58 fish showed sufficient evidence to be conforming.

```{r}
table_lm_1 <- best_mod %>% 
  dplyr::filter(model_type == "lm_1") %>% 
  dplyr::mutate(r_sq_ci = paste0(format(round(r2, 3), scientific = FALSE), " (", 
                   format(round(r2_q2.5, 3), scientific = FALSE), " to ", 
                   format(round(r2_q97.5, 3), scientific = FALSE), ")"),
  est_ci = paste0(format(round(estimate_DO, 4), scientific = FALSE), " (", 
                  format(round(conf.low_DO, 4), scientific = FALSE), " to ", 
                  format(round(conf.high_DO, 4), scientific = FALSE), ")"),
                conformer = if_else(conf.low_DO > 0, "Conforming", "Not conforming")) %>% 
  dplyr::select(id, estimate_DO, r_sq_ci, est_ci, conformer) %>% 
  dplyr::arrange(conformer, desc(est_ci)) %>% 
  dplyr::ungroup()

first_order_ids <- table_lm_1 %>% 
  distinct(id) %>% 
  pull(.)

mean_mo2 <- slope_tidy %>% 
  dplyr::filter(id %in% first_order_ids) %>% 
  dplyr::group_by(id) %>% 
  dplyr::reframe(mean_MO2_g = mean(MO2_g, na.rm = TRUE))

table_lm_1 <- left_join(table_lm_1, mean_mo2, by = "id") 

table_lm_1 <-  table_lm_1 %>% 
  dplyr::mutate(percent_change = (estimate_DO/mean_MO2_g)*100) %>% 
  dplyr::select(id, r_sq_ci, est_ci, percent_change, conformer)


table_lm_1 %>% 
  gt() %>% 
  cols_align(
    align = "center", 
    columns = everything()
  ) %>% 
  cols_label(
    id = "Fish ID",
    r_sq_ci = "r2 (CI)",
    est_ci = "Estimate (CI)",
    percent_change = "Percentage change",
    conformer = "Evidence of oxyconforming"
  )
```

#### Ploting all models

Data set with all slopes and which model was best

```{r}
best_fit <- left_join(slope_tidy, best_mod, by = "id")
```

Saving all regression, and highlighting the model that has the best fit, based on AIC values. This is called `combined_reg_plots.pdf`

```{r, fig.height = 3, fig.width= 4}
# # Create a list to store the plots
# plots <- list()
# model_preds_list <- list()
# incremental_reg_bayes_wd <- paste0(output_fig_wd, "./incremental_regressions")
# 
# for (id_i in ids) {
#   
#   # Filter data for the current ID
#   df_i <- bayes_reg_mods_predictions %>%
#     dplyr::filter(id == id_i) %>% 
#     dplyr::mutate(line_size = if_else(elpd_loo_rank == 1, 2, 1),
#            alpha_value = if_else(elpd_loo_rank == 1, 1, 0.4))
#   
#   x_min <- df_i %>%
#     dplyr::reframe(min = min(DO), na.rm = TRUE) %>% 
#     dplyr::pull(min)
#   
#   y_max <- df_i %>%
#     dplyr::reframe(max = max(MO2_g), na.rm = TRUE) %>% 
#     dplyr::pull(max)
#   
#   best_weighted_model_i <- best_fit_bayes_reg %>% 
#     dplyr::filter(id == id_i & elpd_loo_rank == 1)
#   
#   poly_i_name <- best_weighted_model_i %>%
#     dplyr::mutate(name = case_when(
#       model_type == "lm_0" ~ "0th-order",
#       model_type == "lm_1" ~ "1st-order",
#       model_type == "lm_2" ~ "2nd-order",
#       model_type == "lm_3" ~ "3rd-order",
#       TRUE ~ "ERROR"
#     )) %>% 
#     dplyr::pull(name)
#   
#   r_i <- best_weighted_model_i %>% 
#     dplyr::mutate(r_sq_ci = paste0(round(r2, 3), " (", 
#                                     round(r2_q2.5, 3), "–", 
#                                     round(r2_q97.5, 3), ")")) %>% 
#     dplyr::pull(r_sq_ci)
# 
#   # Create the plot
#   plot_i <- ggplot() +
#     geom_ribbon(data = df_i,
#                 aes(x = DO, y = predicted, ymin = pred_lower, ymax = pred_upper, fill = model_type),
#                 alpha = 0.1) +
#     geom_line(data = df_i, 
#               aes(x = DO, y = predicted, colour = model_type, linewidth = line_size, alpha = alpha_value)) +
#     geom_point(data = df_i %>% dplyr::filter(elpd_loo_rank == 1), aes(x = DO, y = MO2_g), alpha = 0.6, colour = "black", size = 2) +
#     scale_colour_manual(values = c("red", "blue", "green", "purple"), 
#                         labels = c("0th Order", "1st Order", "2nd Order", "3rd Order")) +
#     scale_size_identity() +  # Use the size values directly
#     scale_alpha_identity(guide = "none") +  # Remove the alpha legend 
#     annotate("text", x = x_min, 
#              y = y_max, 
#              label = paste0("Best fit: ",poly_i_name, "\n", "r2 = ", r_i), 
#              hjust = 0, vjust = 1, size = 4) +
#     labs(
#       title = paste(id_i),
#       x = "Dissolved oxygen percentage (DO)",
#       y = "MO2 (o2 mg/g/h)",
#       colour = "Model") +
#     theme_classic() +
#     theme(legend.position = "none")
#   
#   # Store the plot
#   plots[[id_i]] <- plot_i
#   
#   print(plot_i)
# }
# 
# 
# pdf(file = paste0(incremental_reg_bayes_wd, "./combined_reg_plots.pdf"), width = 8, height = 6)
# 
# dev.off()  # Close the PDF device
```

Getting a plot for each best fit regression, and overlaying a global model for that model type.

```{r, echo=FALSE}
output_mods_bayes_global_wd <- paste0(output_mods_wd, "./bayes-regs-global")
```

Here we are grouping fish by best fitting model and getting an average trend. I have hashed the code so you don't need to re-run the models

⏭️ **Skip this step** if you have downloaded the global models **'.rds'** in the 'bayes-regs-global' folder

```{r}
# seed_number = 143019
# 
# # Set up parallel backend
# future::plan(multisession, workers = 4)  # Adjust workers based on system resources
# 
# # Define model formulas and data
# model_formulas <- list(
#   lm_0 = bf(MO2_g ~ 1, family = gaussian()),
#   lm_1 = bf(MO2_g ~ DO, family = gaussian()),
#   lm_2 = bf(MO2_g ~ poly(DO, 2), family = gaussian()),
#   lm_3 = bf(MO2_g ~ poly(DO, 3), family = gaussian())
# )
# 
# model_data <- list(
#   lm_0 = best_fit %>% filter(model_type == "lm_0"),
#   lm_1 = best_fit %>% filter(model_type == "lm_1"),
#   lm_2 = best_fit %>% filter(model_type == "lm_2"),
#   lm_3 = best_fit %>% filter(model_type == "lm_3")
# )
# 
# # Run models in parallel with future_map()
# models <- future_map(names(model_formulas), ~{
#   brm(
#     formula = model_formulas[[.x]],
#     data = model_data[[.x]],
#     cores = 4,
#     seed = seed_number,
#     save_pars = save_pars(all = TRUE),
#     sample_prior = FALSE,
#     silent = TRUE,
#     file = paste0(output_mods_bayes_global_wd, "./", .x, "_global")
#   )
# })
# 
# # Assign model names to results
# names(models) <- names(model_formulas)
# 
# # Stop parallel plan after execution
# future::plan(sequential)
```

⏭️ **Skip this step** if you have downloaded the global_models_pred_df.csv

```{r}
# global_models_preds <- list()
# 
# for (mode_name in names(bayes_reg_mods_global)) {
#   
#   mod_i <- bayes_reg_mods_global[[mode_name]]
#   mode_type = str_remove(mode_name, "_global")
#   
#    model_predictions_i <- as.data.frame(fitted(mod_i, summary = TRUE)) %>% 
#       dplyr::mutate(model = mode_name,
#                     model_type = mode_type) %>% 
#       dplyr::rename(pred_lower = Q2.5, pred_upper = Q97.5, predicted = Estimate, pred_error = Est.Error) %>% 
#       dplyr::select(model, everything())
#    
#    original_data_i <- best_fit %>% filter(model_type == mode_type) %>% 
#      dplyr::select(-model_type)
#     
#     model_predictions_original_i <- cbind(model_predictions_i, original_data_i)
#     
#    
#    global_models_preds[[mode_name]] <- model_predictions_original_i
# }
# 
# global_models_pred_df <- bind_rows(global_models_preds)
# 
# write.csv(global_models_pred_df,  paste0(mod_data_wd, "/.global_models_pred_df.csv"), row.names = FALSE)
```

Load data frame

```{r}
global_models_pred_df <- read.csv(paste0(mod_data_wd, "/.global_models_pred_df.csv"))
```

#### Figure S5

Best fit regressions with global models based on that polynomial order.

```{r}
global_models_pred_df <- global_models_pred_df %>%
  dplyr::mutate(model_name = case_when(
      model_type == "lm_0" ~ "0th-order polynomial",
      model_type == "lm_1" ~ "1st-order polynomial",
      model_type == "lm_2" ~ "2nd-order polynomial",
      model_type == "lm_3" ~ "3rd-order polynomial",
      TRUE ~ "ERROR"
    ))

annotation_data <- mod_summary_table %>%
  dplyr::select(model_type, best_model_name, n)

bayes_reg_mods_predictions_best <- bayes_reg_mods_predictions %>% 
  dplyr::filter(elpd_loo_rank == 1) 

fig_s5 <- ggplot() +
  geom_line(data = bayes_reg_mods_predictions_best,
            aes(x = DO, y = predicted, color = id), size = 1) +
  # geom_smooth(data = best_fit,
  #           aes(x = DO, y = MO2_g, color = id), size = 1, alpha = 1, se = FALSE) +
  geom_ribbon(data = global_models_pred_df,
              aes(x = DO, ymin = pred_lower
, ymax = pred_upper, group = model),
              fill = "#FC6C85", alpha = 0.2) +  # Shaded confidence intervals
  geom_line(data = global_models_pred_df,
            aes(x = DO, y = predicted), linewidth = 1.5, color = "#FF007F") +
  facet_wrap(~model_type) +
  scale_color_grey(start = 0.1, end = 0.9) +
  labs(
      title = paste("Best fit regressions grouped by polynomial order"),
      x = "Dissolved oxygen percentage (DO)",
      y = "MO2 (O2 mg/g/h)") +
  theme_classic() +
  theme(legend.position = "none") +
  geom_text(data = annotation_data,
            aes(x = -Inf, y = Inf, label = paste0("italic(n) == ", n)),
            hjust = -0.1, vjust = 1.2, inherit.aes = FALSE, parse = TRUE)
fig_s5
```

**Figure S5:** Individual regression curves showing the relationship between mass-specific oxygen consumption rate (ṀO₂; mg⁻¹ O₂ h⁻¹) and ambient dissolved oxygen (DO; % saturation), grouped by the best-fitting polynomial order (0ᵗʰ to 3ʳᵈ). Each panel represents fish whose ṀO₂–DO relationship was best modelled by that polynomial order, based on leave-one-out cross-validation (LOO-CV). Grey lines represent individual model fits; the bold pink line shows the group-level trend within each polynomial class. Fish were most commonly best fit by a 2ⁿᵈ-order polynomial (n = 22), followed by 3ʳᵈ-order (*n* = 15), 1ˢᵗ-order (*n* = 11), and 0ᵗʰ-order (**n = 10). Higher-order fits (2ⁿᵈ and 3ʳᵈ) and 0ᵗʰ-order fits are generally indicative of oxyregulating responses, while 1ˢᵗ-order fits may suggest oxyconforming behaviour when positively sloped.  


```{r}
# ggsave(filename = paste0(output_fig_wd, "./ms_figures./figure-2.pdf"), 
#        plot = fig_2,
#        width = 210,
#        height = 297/2,  # Specify the height of the plot
#        units = "mm")
```

## O₂crit model

We will calculate O₂crit using Chabot method and function `calcO2crit`.  

This function uses the fifth percentile of the MO<sub>2</sub> values observed at dissolved oxygen levels ≥ 80% air saturation as the criterion to assess low MO<sub>2</sub> values. The algorithm then identifies all the MO<sub>2</sub> measurements greater than this minimally acceptable MO<sub>2</sub> value. Within this sub-set, it identifies the MO<sub>2</sub>2 measurement made at the lowest DO and thereafter considers this DO as candidate for breakpoint (named pivotDO in the script). A regression is then calculated using observations at DO levels \< pivotDO, and a first estimate of O<sub>2crit</sub> is calculated as the intersection of this regression line with the horizontal line representing SMR. The script then goes through validation steps to ensure that the slope of the regression is not so low that the line, projected to normoxic DO levels, passes below any MO<sub>2</sub> values observed in normoxia. It also ensures that the intercept is not greater than zero. Corrective measures are taken if such problems are encountered.  

lowestMO2 default is the `quantile(MO2[DO >= 80], p=0.05)`. It is used to segment the data and locate the pivotDO.  

```{r}
ids <- slope_tidy %>% 
  dplyr::distinct(id) %>% 
  dplyr::pull()

pcrit_model_df_list <- list()
pcrit_models <-  list()

for (id_i in ids) {

df_i <- slope_tidy %>% 
  dplyr::filter(id == id_i)

o2crit <- calcO2crit(Data = df_i, SMR = df_i$SMR[1], lowestMO2=NA, gapLimit = 4,
                     max.nb.MO2.for.reg = 7)

best_mod_type <- best_mod %>%
  dplyr::filter(id == id_i) %>% 
  pull(model_type)
  
vaule <- o2crit$o2crit
lowestMO2 = quantile(df_i$MO2[df_i$DO >= 80], p=0.05)
SMR <- o2crit$SMR
nb_mo2_conforming <- o2crit$Nb_MO2_conforming
r2 <- o2crit$r2
method <- o2crit$Method
p <- o2crit$P[1]
lethal_point <- min(o2crit$lethalPoints)

pcrit_model_df <- tibble(
      id = id_i,
      pcrit_vaule = vaule,
      pcrit_smr = SMR,
      pcrit_lowestMO2 = lowestMO2,
      pcrit_nb_mo2_conforming = nb_mo2_conforming,
      pcrit_lethal_points = lethal_point,
      pcrit_r2 = r2,
      pcrit_method = method,
      pcrit_p = p,
      best_mod_type = best_mod_type
    )

pcrit_model_df_list[[id_i]] <- pcrit_model_df

pcrit_models[[id_i]] <- o2crit

}

pcrit_model_df <- bind_rows(pcrit_model_df_list)
```

### Ploting O2crit

Here's we are saving all O<sub>2crit</sub> models, even those MO<sub>2</sub>-O<sub>2</sub> relationships best modelled by 0^th^- and 1^st^-order polynomials.

```{r}
# Create output directory if needed
output_fig_pcrit_chabot_wd <- file.path(output_fig_wd, "model_chabot")
if (!dir.exists(output_fig_pcrit_chabot_wd)) {
  dir.create(output_fig_pcrit_chabot_wd)
}

ids <- slope_tidy %>%
  dplyr::distinct(id) %>%
  dplyr::pull()

pcrit_chabot_list <- list()

# Open a single PDF device
pdf(file = file.path(output_fig_pcrit_chabot_wd, "combined_chabot_plots.pdf"),
    width = 8, height = 6)

for (id_i in ids) {

  best_mod_type <- best_mod %>%
  dplyr::filter(id == id_i) %>%
  pull(model_type)

  r2 <- pcrit_model_df %>%
    dplyr::filter(id == id_i) %>%
    dplyr::mutate(pcrit_r2 = round(pcrit_r2, 3)) %>%
    dplyr::pull(pcrit_r2)

  conforming <- pcrit_model_df %>%
    dplyr::filter(id == id_i) %>%
    dplyr::mutate(pcrit_nb_mo2_conforming = round(pcrit_nb_mo2_conforming, 3)) %>%
    dplyr::pull(pcrit_nb_mo2_conforming)

  P <- pcrit_model_df %>%
    dplyr::filter(id == id_i) %>%
    dplyr::mutate(pcrit_p = round(pcrit_p, 3)) %>%
    dplyr::pull(pcrit_p)

  SMR <- pcrit_model_df %>%
    dplyr::filter(id == id_i) %>%
    dplyr::mutate(pcrit_smr = round(pcrit_smr, 3)) %>%
    dplyr::pull(pcrit_smr)

  lowestMO2 <- pcrit_model_df %>%
    dplyr::filter(id == id_i) %>%
    dplyr::mutate(pcrit_lowestMO2 = round(pcrit_lowestMO2, 3)) %>%
    dplyr::pull(pcrit_lowestMO2)

  # Generate and render the plot
  plotO2crit(o2critobj = pcrit_models[[id_i]])

  # Add a title
  mtext(
    text = paste0(id_i),
    side = 3, line = 2, adj = 0,
    col = "blue", font = 2, cex = 1.2
  )

  mtext(
    text = paste0("R2 = ", r2, "; p = ", P, "; CP < SMR = ", conforming, "; SMR = ", SMR, "; lowestMO2 = ",lowestMO2),
    side = 3, line = 1, adj = 0,
    col = "blue", font = 1, cex = 0.8
  )

  mtext(
    text = paste0("Best model fit = ", best_mod_type),
    side = 3, line = 0, adj = 0,
    col = "red", font = 1, cex = 0.8
  )
}

# Close the PDF device *after* the loop
dev.off()
```

Now printing the fish that were fitted with higher order-polynomials in the htlm document

```{r, fig.height = 3, fig.width= 8}

par(mfrow = c(1, 2), mar = c(4, 4, 4, 2)) 

pcrit_model_meta <- slope_tidy %>% 
  dplyr::group_by(id) %>% 
  dplyr::slice(1) %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(pcrit_model_df, ., by = "id")

higher_order_ids <-  pcrit_model_meta %>% 
  dplyr::filter(best_mod_type == "lm_2" | best_mod_type == "lm_3") %>% 
  dplyr::distinct(id) %>% 
  dplyr::pull()

for (id_i in higher_order_ids) {
  
  
  r2 <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_r2 = round(pcrit_r2, 3)) %>% 
    dplyr::pull(pcrit_r2)
  
  conforming <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_nb_mo2_conforming = round(pcrit_nb_mo2_conforming, 3)) %>% 
    dplyr::pull(pcrit_nb_mo2_conforming)
  
  P <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_p = round(pcrit_p, 3)) %>% 
    dplyr::pull(pcrit_p)
  
  SMR <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_smr = round(pcrit_smr, 3)) %>% 
    dplyr::pull(pcrit_smr)
  
  lowestMO2 <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_lowestMO2 = round(pcrit_lowestMO2, 3)) %>% 
    dplyr::pull(pcrit_lowestMO2)
  
  # Generate and render the plot
  plotO2crit(o2critobj = pcrit_models[[id_i]])
  
  # Add a title
  mtext(
    text = paste0(id_i),
    side = 3, line = 2, adj = 0,
    col = "blue", font = 2, cex = 1.2
  )
  
  mtext(
    text = paste0("R2 = ", r2, "; p = ", P, "; CP < SMR = ", conforming),
    side = 3, line = 1, adj = 0,
    col = "blue", font = 0.5, cex = 0.8
  )
  
   mtext(
    text = paste0("SMR = ", SMR, "; lowestMO2 = ",lowestMO2),
    side = 3, line = 0, adj = 0,
    col = "blue", font = 0.5, cex = 0.8
  )
}
```


### O₂crit numerical rules

We applied a numerical rule to asses if a O<sub>2crit</sub> was observed. Specifically, we filtered for only cases were at the lowest O₂ value three consecutive MO₂ measures full below our SMR and fifth percentile of the MO2 values observed at dissolved O2 levels > 80%. In the model output these are called nb_mo2_conforming points.

```{r}
o2crit_numerical_yes <- pcrit_model_df %>% 
  dplyr::filter(pcrit_nb_mo2_conforming > 2) %>% 
  pull(id)

paste0("Based on this rule there are ", length(o2crit_numerical_yes), " fish with possible O₂crits.")
```


#### Figure S6

Plotting fish that meet our numerical rule for O<sub>2crit</sub> 

```{r, fig.height = 3, fig.width= 8}

par(mfrow = c(1, 2), mar = c(4, 4, 4, 2)) 

for (id_i in o2crit_numerical_yes) {
  
  
  r2 <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_r2 = round(pcrit_r2, 3)) %>% 
    dplyr::pull(pcrit_r2)
  
  conforming <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_nb_mo2_conforming = round(pcrit_nb_mo2_conforming, 3)) %>% 
    dplyr::pull(pcrit_nb_mo2_conforming)
  
  P <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_p = round(pcrit_p, 3)) %>% 
    dplyr::pull(pcrit_p)
  
  SMR <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_smr = round(pcrit_smr, 3)) %>% 
    dplyr::pull(pcrit_smr)
  
  lowestMO2 <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_lowestMO2 = round(pcrit_lowestMO2, 3)) %>% 
    dplyr::pull(pcrit_lowestMO2)
  
  # Generate and render the plot
  plotO2crit(o2critobj = pcrit_models[[id_i]])
  
  # Add a title
  mtext(
    text = paste0(id_i),
    side = 3, line = 2, adj = 0,
    col = "blue", font = 2, cex = 1.2
  )
  
  mtext(
    text = paste0("R2 = ", r2, "; p = ", P, "; CP < SMR = ", conforming),
    side = 3, line = 1, adj = 0,
    col = "blue", font = 0.5, cex = 0.8
  )
  
   mtext(
    text = paste0("SMR = ", SMR, "; lowestMO2 = ",lowestMO2),
    side = 3, line = 0, adj = 0,
    col = "blue", font = 0.5, cex = 0.8
  )
}
```

**Figure S6**: Representative examples of fish that were best modelled by a 2nd- or 3rd-order polynomial and meet the numerical inclusion criteria. The horizontal solid orange line shows the standard metabolic rate (SMR), the horizontal solid pink line shows thelowestMO2 (as defined by  Claireaux and Chabot (2016)), the solid/dashed blue/green slope line shows the O2 regulation failure regression (based on the rule-based linear regression method of Claireaux and Chabot (2016); where the colour and line type indicate the type of regression fitted, through orgin or not), and the solid vertical blue/green line shows the O2crit estimate (intersection of O2 failure regression and SMR).

How many from each salinity treatment

```{r}
slope_tidy %>% 
  dplyr::filter(id %in% o2crit_numerical_yes) %>% 
  dplyr::group_by(id) %>% 
  dplyr::slice(1) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe(n = length (id)) %>% 
  dplyr::ungroup() %>% 
  gt()
```

### Visual inspection

We conducted visual inspection of all 58 fish to determine if O<sub>2crits</sub> were present. We have loaded this as a data frame **💿 o2crit_check**

For this visual inspection, we grouped fish into "no", "maybe" or "yes", based on the certainty at which we observed a O<sub>2crits</sub>. In the case of yes or maybe, we also estimate dissolved oxygen percentage. All authors did this independently.  

#### 📈 Results

Let's figure out how consistent the categorisation was, and what the majority response was. There were 35 cases with no disagreements (60.34%), 23 cases with one or more disagreements (39.66%).  

```{r}
disagreement_summary <- o2crit_check %>%
  group_by(id) %>%
  dplyr::reframe(
    n_disagreement = n_distinct(pcrit_cat)-1,
    cat_disagreement = if_else(n_disagreement==0, 0, 1),
    majority_cat = names(sort(table(pcrit_cat), decreasing = TRUE))[1],
    certainty = max(table(pcrit_cat)/5*100)
  )

total_assesment <- o2crit_check %>% 
  distinct(id) %>% 
  nrow()

disagreement_summary %>%
  group_by(cat_disagreement) %>% 
  dplyr::reframe(
    n = length(id),
    percent = round(n/total_assesment*100,2)
  ) %>% 
  gt()
```

#### Figure S7

Here's a heat map to show decisions for each fish

```{r, fig.height=10, fig.width=5}
# Define custom colours for pcrit_cat
category_colours <- c(
  "no" = "#E03C32",
  "maybe" = "#FFD301",
  "yes" = "#7BB662"
)

# Create heatmap
ggplot(o2crit_check, aes(x = expert, y = id, fill = pcrit_cat)) +
  geom_tile(colour = "white") +
  scale_fill_manual(values = category_colours, na.value = "grey90") +
  labs(
    title = "Visual Assessment of pcrit_cat by Expert and ID",
    x = "Expert",
    y = "Fish ID",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 7), # adjust size as needed
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )
```


**Figure S7**: Heatmap showing the visual assessment of critical oxygen saturation (O<sub>2crits</sub>) category ("no", "maybe" or "yes") for each fish by expert assessor (Elizabeth C. Hoots [ech], Jake M. Martin [jmm], Luis L. Kuchenmülle [llk] Maryane Gradito [mg], Timothy D. Clark [tdc]). Each tile represents one fish-expert pairing, colour-coded by assigned category:
<span style="color:#E03C32"><strong>no</strong></span> (red, #E03C32),
<span style="color:#FFD301"><strong>maybe</strong></span> (yellow, #FFD301), and
<span style="color:#7BB662"><strong>yes</strong></span> (green, #7BB662).

Expert deviation from the majority assessment ranged from 22.41 to 6.90%.

```{r}
o2crit_check_sum <- o2crit_check %>% 
  dplyr::left_join(., disagreement_summary, by = "id") %>% 
  dplyr::mutate(deviated = if_else(pcrit_cat != majority_cat, 1, 0))

o2crit_check_sum %>%
  group_by(expert) %>%
  summarise(
    deviations = sum(deviated),
    deviation_rate = round(deviations / 58 * 100, 2)
  ) %>% 
  dplyr::arrange(desc(deviations)) %>% 
  gt()
```

Let's make a list of those that had majority assigned as yes for an O<sub>2crit</sub>. There was 10 total.  

```{r}
o2crit_visual_yes <-  disagreement_summary %>% 
  dplyr::filter(majority_cat == "yes") %>% 
  dplyr::pull(id)

paste0("Based on the majority visual assessment ", length(o2crit_visual_yes), " fish have possible O₂crit.")
```

### O₂crit numbers

We will now check to see which fish have O<sub>2crit</sub> based on the numerical and visual inspections.   

There are eight fish that both numerically and visually were determined to have a O<sub>2crit</sub>. Two fish with a visually confirmed O<sub>2crit</sub>, that did not meet the numerical criteria. Eight fish meet the numerical criteria, but did not pass the visual inspection.  

```{r}
shared_elements <- intersect(o2crit_visual_yes, o2crit_numerical_yes)

unique_to_visual <- setdiff(o2crit_visual_yes, o2crit_numerical_yes)

unique_to_numerical <- setdiff(o2crit_numerical_yes, o2crit_visual_yes)

list(
  Shared = shared_elements,
  Unique_to_visual = unique_to_visual,
  Unique_to_numerical = unique_to_numerical
)
```

How many per salinity treatment 

```{r}
slope_tidy %>% 
  dplyr::filter(id %in% shared_elements) %>% 
  dplyr::group_by(id) %>% 
  dplyr::slice(1) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe(n = length (id)) %>% 
  dplyr::ungroup() %>% 
  gt()
```


#### 📈 Results

Looking at mean O<sub>2crit</sub> estimates.   

```{r}
pop_mass <- pcrit_model_meta %>% 
  dplyr::reframe(
    mean_mass = mean(mass)
  ) %>% 
  pull(mean_mass)

pcrit_model_meta %>% 
  dplyr::filter(id %in% shared_elements) %>% 
  dplyr::mutate(rel_mass = mass/pop_mass) %>% 
  dplyr::reframe(
    n_o2crit = length(pcrit_vaule),
    mean_o2crit = mean(pcrit_vaule),
    min_o2crit = min(pcrit_vaule),
    max_o2crit = max(pcrit_vaule),
    mean_mass = mean(mass),
    mean_rel_mass = mean (rel_mass),
    sd_mass = sd(mass),
    sd_rel_mass = sd(rel_mass)
  ) %>% 
  gt() %>% 
    fmt_number(
    columns = where(is.numeric),
    decimals = 2
  )
```


```{r}
min_dos <- slope_tidy %>% 
  dplyr::filter(!(id %in% shared_elements)) %>%
  group_by(id) %>% 
  dplyr::reframe(
    min_do = min(DO)
  )

min_dos %>% 
  dplyr::reframe(
    mean_min_do = round(mean(min_do),2),
    sd_min_do = round(sd(min_do),2),
    min_min_do = round(min(min_do),2)
  ) %>% 
  gt()
```




```{r}
pcrit_model_meta %>% 
  dplyr::filter(id %in% shared_elements) %>% 
  dplyr::group_by(salinity_group) %>% 
  dplyr::reframe(
    n_o2crit = length(pcrit_vaule),
    mean_o2crit = mean(pcrit_vaule),
    min_o2crit = min(pcrit_vaule),
    max_o2crit = max(pcrit_vaule),
    mean_mass = mean(mass)
  ) %>% 
  gt()
```

### Figure S8

These are the 8 fish that meet numerical criteria and passed visual inspection.

```{r, fig.height = 3, fig.width= 8}

par(mfrow = c(1, 2), mar = c(4, 4, 4, 2)) 

for (id_i in shared_elements) {
  
  r2 <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_r2 = round(pcrit_r2, 3)) %>% 
    dplyr::pull(pcrit_r2)
  
  conforming <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_nb_mo2_conforming = round(pcrit_nb_mo2_conforming, 3)) %>% 
    dplyr::pull(pcrit_nb_mo2_conforming)
  
  P <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_p = round(pcrit_p, 3)) %>% 
    dplyr::pull(pcrit_p)
  
  SMR <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_smr = round(pcrit_smr, 3)) %>% 
    dplyr::pull(pcrit_smr)
  
  lowestMO2 <- pcrit_model_meta %>% 
    dplyr::filter(id == id_i) %>% 
    dplyr::mutate(pcrit_lowestMO2 = round(pcrit_lowestMO2, 3)) %>% 
    dplyr::pull(pcrit_lowestMO2)
  
  # Generate and render the plot
  plotO2crit(o2critobj = pcrit_models[[id_i]])
  
  # Add a title
  mtext(
    text = paste0(id_i),
    side = 3, line = 2, adj = 0,
    col = "blue", font = 2, cex = 1.2
  )
  
  mtext(
    text = paste0("R2 = ", r2, "; p = ", P, "; CP < SMR = ", conforming, "; SMR = ", SMR, "; lowestMO2 = ",lowestMO2),
    side = 3, line = 1, adj = 0,
    col = "blue", font = 1, cex = 0.8
  )
}
```

**Figure S8**: Representative examples of fish that were best modelled by a 2nd- or 3rd-order polynomial and meet the numerical inclusion criteria. The horizontal solid orange line shows the standard metabolic rate (SMR), the horizontal solid pink line shows thelowestMO2 (as defined by  Claireaux and Chabot (2016)), the solid/dashed blue/green slope line shows the O2 regulation failure regression (based on the rule-based linear regression method of Claireaux and Chabot (2016); where the colour and line type indicate the type of regression fitted, through orgin or not), and the solid vertical blue/green line shows the O2crit estimate (intersection of O2 failure regression and SMR).

Pulling model method and lethal points

```{r}
ids <- pcrit_model_meta %>% 
  dplyr::distinct(id) %>% 
  dplyr::pull()

pcrit_mean_plot <- list()

for (id_i in ids) {
  best_mod_type <- best_mod %>%
    dplyr::filter(id == id_i) %>% 
    pull(model_type)
  mod <- pcrit_models[[id_i]]
  data <- mod$origData
  min_leathal <- min(mod$lethalPoints)
  method <- mod$Method
  pcrit <- mod$o2crit
  data_slope <- data %>%
    dplyr::mutate(n_row = 1:nrow(.),
                  leathal = if_else(n_row >= min_leathal, "nonregulation", "regulation"),
                  method = method,
                  pcrit = pcrit,
                  best_mod_type = best_mod_type
    )
  pcrit_mean_plot[[id_i]] <- data_slope
}

pcrit_mean_plot_df <- bind_rows(pcrit_mean_plot)
```


### Figure S9

Making a O2crit plot for each of the eight fish that past visual and  numerical assessment on the same axis.   

```{r}
pcrit_summary <- pcrit_mean_plot_df %>%
  dplyr::filter(id %in% shared_elements) %>% 
  group_by(id) %>%
  summarise(
    SMR = first(SMR),
    pcrit = first(pcrit),
    max_DO = max(DO, na.rm = TRUE),
    method = first(method),
    .groups = "drop"
  )

smooth_segments <- map_dfr(unique(pcrit_summary$id), function(cur_id) {
  
  sum_info <- pcrit_summary %>% filter(id == cur_id)
  SMR_val <- sum_info$SMR
  method_val <- sum_info$method
  
  df <- pcrit_mean_plot_df %>%
    filter(id == cur_id, leathal == "nonregulation")
  
  if(nrow(df) < 2) return(NULL)
  
  # Fit the appropriate model
  if(method_val == "through_origin") {
    model <- lm(MO2 ~ DO + 0, data = df)
  } else {
    model <- lm(MO2 ~ DO, data = df)
  }
  
  coefs <- coef(model)
  
  if(method_val == "through_origin") {
    slope <- coefs[1]
    DO_start <- 0
    DO_stop <- SMR_val / slope
  } else {
    intercept <- coefs[1]
    slope <- coefs[2]
    # For LS_reg, start at 0 if prediction is already positive,
    # or at the DO where predicted MO2 becomes 0 if not.
    DO_start <- if(intercept < 0) -intercept / slope else 0
    DO_stop <- (SMR_val - intercept) / slope
  }
  
  # IMPORTANT: Do not bound DO_stop by max(df$DO) if you want to extrapolate
  # DO_stop <- min(DO_stop, max(df$DO, na.rm = TRUE))  # <-- Remove or comment out this line
  
  # Create a sequence from DO_start to DO_stop
  DO_seq <- seq(DO_start, DO_stop, length.out = 100)
  
  # Calculate the predicted MO2 values for this sequence
  MO2_pred <- if(method_val == "through_origin") {
    slope * DO_seq
  } else {
    intercept + slope * DO_seq
  }
  
  tibble(
    id = cur_id,
    DO = DO_seq,
    MO2_pred = MO2_pred
  ) 
  })

Fig_s9 <- ggplot() +
  geom_point(data = pcrit_mean_plot_df %>%
               dplyr::filter(id %in% shared_elements), 
             aes(x = DO, y = MO2, colour = id, alpha = leathal)) +
  scale_alpha_manual(values = c("regulation" = 0.2, "nonregulation" = 0.8)) +
  guides(alpha = "none") +
  geom_line(data = smooth_segments,
            aes(x = DO, y = MO2_pred, colour = id), linewidth = 1.2) +
  geom_segment(data = pcrit_summary,
             aes(x = pcrit, xend = max_DO, y = SMR, yend = SMR)) +
  geom_segment(data = pcrit_summary,
               aes(x = pcrit, xend = pcrit, y = 0, yend = SMR)) +
  scale_colour_viridis_d(option = "plasma") +
  theme_classic() +
  labs(
    y = expression("mg O"[2]~"h"^-1),
    x = "Dissolved oxygen (DO; % saturation)"
  ) +
  theme(legend.position = "bottom")

Fig_s9
```


**Figure S9**: Oxygen consumption rate (ṀO₂; mg⁻¹ O₂ h⁻¹) and ambient dissolved oxygen (DO; % saturation) plotted for each fish that passed both past visual and  numerical assessment for O<sub>2crit</sub>. The horizontal dashed lines show SMR, the solid sloped lines represent the linear best fit for ṀO values below regulatory failure. The vertical soild lines represent O<sub>2crit</sub> (the intersection of the linear best fit for ṀO values below regulatory failure and SMR).



### Figure 2

Let's make some example plots for each major types of O2-MO2 relationships (i.e. 0^th^-, 1^st^-, and 2^nd^/3^rd^- order polynomial).   

First an example of a 0th-order polynomial fit (c_9_24nov_4).   

```{r}
fig_2a <- bayes_reg_mods_predictions %>%
  dplyr::filter(id == "c_9_24nov_4", model_type == "lm_0") %>% 
  ggplot() +
  geom_point(aes(x = DO, y = MO2), colour = "black", size = 3, alpha = 0.5) +
  geom_segment(
    aes(x = 0, xend = 100, y = 0, yend = SMR[1]), # SMR
    linetype = "dashed", size = 1, colour = "#0F7AAF") +
  geom_segment(
    aes(x = 0, xend = 100, y = SMR[1], yend = SMR[1]), # conform
    linetype = "dashed", size = 1, colour = "#FFA500") +
  theme_few() +
    labs(
    y = expression("mg O"[2]~"h"^-1),
    x = NULL
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```


Second an example of a 1st-order polynomial best fit (d_9_25nov_3).  

```{r}
fig_2b <- bayes_reg_mods_predictions %>%
  dplyr::filter(id == "d_9_25nov_3", model_type == "lm_1") %>% 
  ggplot() +
  geom_point(aes(x = DO, y = MO2), colour = "black", size = 3, alpha = 0.5) +
  # geom_line(aes(x = DO, y = predicted * mass), size = 2, colour = "blue") +
  geom_segment(
    aes(x = 0, xend = 100, y = 0, yend = SMR[1]), # SMR
    linetype = "dashed", size = 1, colour = "#0F7AAF") +
  geom_segment(
    aes(x = 0, xend = 100, y = SMR[1], yend = SMR[1]), # conform
    linetype = "dashed", size = 1, colour = "#FFA500") +
  theme_few() +
  labs(
    y = expression("mg O"[2]~"h"^-1),
    x = NULL
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

```


Last an example of a fish best fit with a higher order polynomial (2nd-order).

```{r}
# Define the example ID
pcrit_example <- "b_0_24nov_1"

# 1. Prepare summary data for the selected ID
pcrit_summary_example <- pcrit_mean_plot_df %>%
  filter(id == pcrit_example) %>%
  summarise(
    SMR = first(SMR),
    pcrit = first(pcrit),
    max_DO = max(DO, na.rm = TRUE),
    method = first(method),
    .groups = "drop"
  )

# 2. Subset the original nonregulation data
df <- pcrit_mean_plot_df %>%
  filter(id == pcrit_example, leathal == "nonregulation")

# 3. Fit model and generate predictions
if (nrow(df) >= 2) {
  SMR_val <- pcrit_summary_example$SMR
  method_val <- pcrit_summary_example$method
  
  if (method_val == "through_origin") {
    model <- lm(MO2 ~ DO + 0, data = df)
    slope <- coef(model)[1]
    DO_start <- 0
    DO_stop <- SMR_val / slope
    MO2_pred <- slope * seq(DO_start, DO_stop, length.out = 100)
  } else {
    model <- lm(MO2 ~ DO, data = df)
    intercept <- coef(model)[1]
    slope <- coef(model)[2]
    DO_start <- if (intercept < 0) -intercept / slope else 0
    DO_stop <- (SMR_val - intercept) / slope
    DO_seq <- seq(DO_start, DO_stop, length.out = 100)
    MO2_pred <- intercept + slope * DO_seq
  }
  
  DO_seq <- seq(DO_start, DO_stop, length.out = 100)
  
  smooth_segment <- tibble(
    id = pcrit_example,
    DO = DO_seq,
    MO2_pred = MO2_pred
  )
}

# 4. Plot the figure
fig_2c <- ggplot() +
  geom_point(
    data = pcrit_mean_plot_df %>% filter(id == pcrit_example),
    aes(x = DO, y = MO2, colour = leathal), size = 3, alpha = 0.5
  ) +
  scale_colour_manual(values = c("regulation" = "black", "nonregulation" = "#c30010")) +
  
  # geom_smooth(data = pcrit_mean_plot_df %>% filter(id == pcrit_example),
  #             aes(x = DO, y = MO2), method = "lm", formula = y ~ poly(x, 3), se = FALSE, alpha = 0.8, size = 2, fullrange = TRUE) +
  
  geom_line(
    data = smooth_segment,
    aes(x = DO, y = MO2_pred), size = 2, colour = "#c30010" #failure  slope
  ) +
  
  
  geom_segment(
    data = pcrit_summary_example,
    aes(x = 0, xend = max_DO, y = SMR, yend = SMR), #SMR
    colour = "#FFA500", linetype = "dashed", size = 1) +
  
  geom_segment(
    data = pcrit_summary_example,
    aes(x = 0, xend = max_DO, y = 0, yend = SMR), #confrom
    linetype = "dashed", size = 1, colour = "#0F7AAF") +
  
  geom_segment(
    data = pcrit_summary_example,
    aes(x = pcrit, xend = pcrit, y = 0, yend = SMR), colour = "darkgreen", size = 2 #Pcrit
  ) +
  
  #coord_cartesian(ylim = c(0, 0.16)) +
  theme_few() +
   labs(
    y = expression("mg O"[2]~"h"^-1),
    x = "Dissolved oxygen (DO; % saturation)"
  ) +
  theme(legend.position = "none")

```


Let's combined them in one plot

```{r, fig.height=10, fig.width=5}
fig_2_bind <- grid.arrange(fig_2a, fig_2b, fig_2c, ncol = 1) #bind plot
```

**Figure 2**: Figure 2. Representative examples of fish that were best modelled by a 0th-order polynomial (A), 1st-order polynomial (B), and 2nd-order polynomial (C). The orange dashed line shows the SMR, while the blue dashed line shows the theoretical oxyconfroming regression. For plot (C), the red sloped line shows the O2 regulation failure regression (based on the rule-based linear regression method of Claireaux and Chabot, 2016), and the solid horizontal green line shows the O2crit estimate (intersection of O2 failure regression and SMR).  

```{r}
# ggsave(filename = paste0(output_fig_wd, "./ms_figures./figure-2.pdf"),
#        plot = fig_2_bind,
#        width = 210,
#        height = 297,
#        units = "mm")
```


# Comapring to past data

Here, we have recreated a similar plot to that presented in Urbina, Glover, and Forster (2012)^[1]^ and have extracted the mean level data from Figure 1a using the metaDigitise package in R^[3]^. This data is called urbina_et_al_2012. This allows us to compare the differences in Mo2 and the relationship between Mo2 and O<sub>2</sub>.

First making a binned data frame to match Urbina, Glover, and Forster (2012) as closely as possible.

```{r}
min_o2_kpa <- min(slope_tidy$o2_kpa, na.rm = TRUE)
max_o2_kpa <- max(slope_tidy$o2_kpa, na.rm = TRUE)

time_bin_df <- slope_tidy %>%
  mutate(o2_group = cut(o2_kpa, 
                        breaks = seq(min_o2_kpa, max_o2_kpa, length.out = 13), # 11 intervals, so 12 breakpoints
                        labels = paste0("Group ", 1:12), 
                        include.lowest = TRUE)) %>% 
  dplyr::group_by(o2_group) %>% 
  dplyr::reframe(mean_MO2_g = mean(MO2_g)*31.25,
                 mean_o2_kpa = mean(o2_kpa),
                 n = length(MO2_g)*31.25,
                 MO2_g_sd = sd(MO2_g)*31.25,
                 o2_kpa_sd = sd(o2_kpa))
```

## Figure 3

Now the plot with our mean data and the mean data from Urbina, Glover, and Forster (2012)^[1]^   

```{r}
n <- slope_tidy %>% 
  dplyr::distinct(id) %>% 
  nrow(.)

# Existing plot
fig3 <- time_bin_df %>% 
  ggplot() +
  geom_point(data = slope_tidy, 
             aes(y = MO2_g * 31.25, x = o2_kpa), 
             size = 2, color = "grey", alpha = 0.3) +  
  geom_point(data = time_bin_df, 
             aes(y = mean_MO2_g, x = mean_o2_kpa),
             size = 3, colour = "#0E4C92", show.legend = FALSE) +
  geom_errorbar(data = time_bin_df,
                aes(y = mean_MO2_g, x = mean_o2_kpa,
                    ymin = mean_MO2_g - MO2_g_sd, ymax = mean_MO2_g + MO2_g_sd), 
                width = 0.15, colour = "#0E4C92") +
  geom_errorbarh(data = time_bin_df, 
                 aes(y = mean_MO2_g, x = mean_o2_kpa,
                     xmin = mean_o2_kpa - o2_kpa_sd, xmax = mean_o2_kpa + o2_kpa_sd), 
                 height = 0.4, colour = "#0E4C92") +
  geom_point(data = urbina_et_al_2012, 
             aes(x = o2_mean, y = mo2_mean), 
             size = 3, shape = 1, fill = "#D21F3C", color = "#D21F3C", stroke = 1) +
  geom_errorbar(data = urbina_et_al_2012, 
                aes(x = o2_mean, ymin = mo2_mean - mo2_sd, ymax = mo2_mean + mo2_sd), 
                width = 0.2, colour = "#D21F3C") +
  geom_errorbarh(data = urbina_et_al_2012, 
                 aes(y = mo2_mean, xmin = o2_mean - o2_sd, xmax = o2_mean + o2_sd), 
                 height = 0.4, colour = "#D21F3C") +
  annotate("text", x = 0, 
           y = 16, 
           label = bquote(atop("Current data (blue), " * italic(n) * " = " * .(n),
                        "Urbina data (red), " * italic(n) * " = 67")),  
         hjust = 0, vjust = 1, size = 4) +
  theme(
    legend.position = "none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "white", colour = NA),
    plot.background = element_rect(fill = "white", colour = NA)
  ) +
  labs(
  subtitle = NULL,
  x = expression("PO"[2]~"(kPa)"),
  y = expression("MO"[2]~"(μmol O"[2]~"g"^-1~"h"^-1)) +
  scale_y_continuous(limits = c(0, 16), breaks = seq(0, 16, by = 2)) +
  scale_x_continuous(limits = c(0, 22), breaks = seq(0, 22, by = 2))  

fig3
```

**Figure 3:** Mean and standard error of metabolic rate (MO<sub>2</sub> μmol O<sub>2</sub> g^-1^ h^-1^) and oxygen concentration (PO<sub>2</sub> kPa) using 12 evenly spaced bins over the O<sub>2</sub> range of observed data (blue filled dots). Compared against the mean and standard error reported in Urbina (2012)^[1]^ (red open dots). The grey dots are the raw observed data form the present study.

```{r}
# ggsave(filename = paste0(output_fig_wd, "./ms_figures./figure-3.pdf"),
#        plot = fig3,
#        width = 210,
#        height = 297/2,  # Specify the height of the plot
#        units = "mm")
```

<!------------------------------->

# 📚 References

<!------------------------------->

^[1]^ Claireaux, G. and Chabot, D. (2016) Responses by fishes to environmental hypoxia: integration through Fry's concept of aerobic metabolic scope. Journal of Fish Biology <https://doi.org/10.1111/jfb.12833>

^[2]^ Urbina MA, Glover CN, and Forster ME, (2012) A novel oxyconforming response in the freshwater fish *Galaxias maculatus*. Comparative Biochemistry and Physiology Part A: Molecular & Integrative Physiology. <https://doi.org/10.1016/j.cbpa.2011.11.011>

^[3]^ Pick JL, Nakagawa S, and Noble DWA (2018) Reproducible, flexible and high-throughput data extraction from primary literature: The metaDigitise r package. <https://doi.org/10.1111/2041-210X.13118>

<!------------------------------->

# 💻 Session information

<!------------------------------->

```{r}
citation()
```

Here is a detailed list of the session information

```{r}
sessioninfo::session_info()
```
